---
title: "20151201"
author: "Derek Sollberger"
date: "10/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
```

--- Day 2: I Was Told There Would Be No Math ---

The elves are running low on wrapping paper, and so they need to submit an order for more. They have a list of the dimensions (length l, width w, and height h) of each present, and only want to order exactly as much as they need.

Fortunately, every present is a box (a perfect right rectangular prism), which makes calculating the required wrapping paper for each gift a little easier: find the surface area of the box, which is 2*l*w + 2*w*h + 2*h*l. The elves also need a little extra paper for each present: the area of the smallest side.

For example:

    A present with dimensions 2x3x4 requires 2*6 + 2*12 + 2*8 = 52 square feet of wrapping paper plus 6 square feet of slack, for a total of 58 square feet.
    A present with dimensions 1x1x10 requires 2*1 + 2*10 + 2*10 = 42 square feet of wrapping paper plus 1 square foot of slack, for a total of 43 square feet.

All numbers in the elves' list are in feet. How many total square feet of wrapping paper should they order?


Puzzle Input:

```{r}
paper_orders <- read.table("20151202input.txt")
colnames(paper_orders) <- "specs"
```

```{r}
# parse length, width, and height
paper_orders <- paper_orders %>%
  separate(specs,
           into = c("length", "width", "height"),
           sep = "x",
           remove = FALSE)

paper_orders$length <- as.numeric(paper_orders$length)
paper_orders$width <- as.numeric(paper_orders$width)
paper_orders$height <- as.numeric(paper_orders$height)
```
  
```{r}
#find smallest side
paper_orders <- paper_orders %>%
  rowwise() %>%
  # mutate(smallest_dim = min(c_across(length:height)),
  #        median_dim = median(c_across(length:height)))
  mutate(smallest_side = 
         min(c_across(length:height))*median(c_across(length:height)))
```

```{r}
# compute area needed per package
paper_orders <- paper_orders %>%
  mutate(wrapping_area = 2*length*width + 2*width*height + 
           2*length*height + smallest_side)
```

```{r}
print(sum(paper_orders$wrapping_area))
```

