---
title: "Day 5"
author: "Derek Sollberger"
date: "11/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
```

--- Day 5: Doesn't He Have Intern-Elves For This? ---

Santa needs help figuring out which strings in his text file are naughty or nice.

A nice string is one with all of the following properties:

    It contains at least three vowels (aeiou only), like aei, xazegov, or aeiouaeiouaeiou.
    It contains at least one letter that appears twice in a row, like xx, abcdde (dd), or aabbccdd (aa, bb, cc, or dd).
    It does not contain the strings ab, cd, pq, or xy, even if they are part of one of the other requirements.

For example:

    ugknbfddgicrmopn is nice because it has at least three vowels (u...i...o...), a double letter (...dd...), and none of the disallowed substrings.
    aaa is nice because it has at least three vowels and a double letter, even though the letters used by different rules overlap.
    jchzalrnumimnmhp is naughty because it has no double letter.
    haegwjzuvuyypxyu is naughty because it contains the string xy.
    dvszwmarrgswjxmb is naughty because it contains only one vowel.

How many strings are nice?

```{r}
# load data
santas_list_raw <- read.table("20151205input.txt")
```

```{r}
# helper function
vowel_check <- function(child_name){
  pass_check <- FALSE #initialization
  
  if(str_count(child_name, "a") + str_count(child_name, "e") + 
     str_count(child_name, "i") + str_count(child_name, "o") + 
     str_count(child_name, "u") >= 3){ pass_check <- TRUE}
  
  # return
  pass_check
}

# vowel_check("ugknbfddgicrmopn")
# vowel_check("dvszwmarrgswjxmb")
```

```{r}
# helper function
doubles_check <- function(child_name){
  pass_check <- FALSE
  child_name_list <- stringr::str_split(child_name, "")[[1]]
  n <- length(child_name_list)
  
  for(i in 1:(n-1)){
    if(child_name_list[i] == child_name_list[i+1]){
      pass_check <- TRUE
      break
    }
  }
  
  #return
  pass_check
}

# doubles_check("ugknbfddgicrmopn")
# doubles_check("jchzalrnumimnmhp")
```

```{r}
# helper function
red_flag_check <- function(child_name){
  pass_check <- TRUE
  
  if(str_count(child_name, "ab") > 0){ pass_check <- FALSE }
  if(str_count(child_name, "cd") > 0){ pass_check <- FALSE }
  if(str_count(child_name, "pq") > 0){ pass_check <- FALSE }
  if(str_count(child_name, "xy") > 0){ pass_check <- FALSE }
  
  #return
  pass_check
}

# red_flag_check("ugknbfddgicrmopn")
# red_flag_check("haegwjzuvuyypxyu")
```

```{r}
# main loop
nice_children <- 0 #initialization
N <- nrow(santas_list_raw)

for(i in 1:N){
  if(vowel_check(santas_list_raw[i,1])){
    if(doubles_check(santas_list_raw[i,1])){
      if(red_flag_check(santas_list_raw[i,1])){
        nice_children <- nice_children + 1
      }
    }
  }
}

print(paste0("There are ", nice_children, " nice children"))
```

