---
title: "Day 4"
author: "Derek Sollberger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
```

--- Day 4: High-Entropy Passphrases ---

A new system policy has been put in place that requires all accounts to use a passphrase instead of simply a password. A passphrase consists of a series of words (lowercase letters) separated by spaces.

To ensure security, a valid passphrase must contain no duplicate words.

For example:

    aa bb cc dd ee is valid.
    aa bb cc dd aa is not valid - the word aa appears more than once.
    aa bb cc dd aaa is valid - aa and aaa count as different words.

The system's full passphrase list is available as your puzzle input. How many passphrases are valid?

```{r}
puzzle_input <- readLines("20171204input.txt")
N <- length(puzzle_input)
```

```{r}
is_valid <- function(this_string){
  pass_check <- TRUE
  this_split <- str_split(this_string, " ")[[1]]
  n <- length(this_split)
  for(j in 1:(n-1)){
    for(k in 2:n){
      if(j != k){
        if(this_split[j] == this_split[k]){
          pass_check <- FALSE
        }
      }
    }
  }
  pass_check #return
}
```


```{r}
num_valid <- 0
for(i in 1:N){
  if(is_valid(puzzle_input[i])){
    num_valid <- num_valid + 1
  }
}
print(paste("There are", num_valid, "valid passphrases"))
# 166 is too low
# 451 was the correct answer
```
--- Part Two ---

For added security, yet another system policy has been put in place. Now, a valid passphrase must contain no two words that are anagrams of each other - that is, a passphrase is invalid if any word's letters can be rearranged to form any other word in the passphrase.

For example:

    abcde fghij is a valid passphrase.
    abcde xyz ecdab is not valid - the letters from the third word can be rearranged to form the first word.
    a ab abc abd abf abj is a valid passphrase, because all letters need to be used when forming another word.
    iiii oiii ooii oooi oooo is valid.
    oiii ioii iioi iiio is not valid - any of these words can be rearranged to form any other word.

Under this new system policy, how many passphrases are valid?

```{r}
is_valid <- function(this_string){
  pass_check <- TRUE
  this_split <- str_split(this_string, " ")[[1]]
  n <- length(this_split)
  for(j in 1:(n-1)){
    for(k in 2:n){
      if(j != k){
        word1 <- this_split[j]
        word2 <- this_split[k]
        anagram1 <- sort(str_split(word1, "")[[1]])
        anagram2 <- sort(str_split(word2, "")[[1]])
        
        if( identical(anagram1, anagram2) ){
          pass_check <- FALSE
        }
      }
    }
  }
  pass_check #return
}
```

```{r}
num_valid <- 0
for(i in 1:N){
  if(is_valid(puzzle_input[i])){
    num_valid <- num_valid + 1
  }
}
print(paste("There are", num_valid, "valid passphrases"))
# 223 was the correct answer!
```