---
title: "Day 11"
author: "Derek Sollberger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
```

```{r}
t1 <- proc.time()
puzzle_input <- c(4438624262, 6263251864, 2618812434, 2134264565, 1815131247, 2612457325,
8585767584, 7217134556, 2825456563, 8248473584)
#example
# puzzle_input <- c(5483143223, 2745854711, 5264556173, 6141336146, 6357385478, 4167524645, 2176841721, 6882881134, 4846848554, 5283751526)
m <- length(puzzle_input) #number of rows
n <- length(str_split(as.character(puzzle_input[1]), "")[[1]]) #number of columns

# create a matrix (but with a buffer on all 4 sides)
M <- matrix(0, nrow = m + 2, ncol = n + 2)
for(row in 2:(m+1)){
  M[row,2:(n+1)] <- as.integer(str_split(as.character(puzzle_input[row-1]), "")[[1]])
}

# simulate the octupuses flashes
num_flashes <- 0
M_flashed <- matrix(FALSE, nrow = m + 2, ncol = n + 2) #keep track of which octopuses flashed
for(timestep in 1:100){ 
  
  #add one energy unit each octopus (avoid buffer in first and last rows and columns)
  M[2:(m+1),2:(n+1)] <- M[2:(m+1),2:(n+1)] + 1
  
  for(k in 1:(m*n)){ #avoid while loop
    #traverse matrix (avoid buffer in first and last rows and columns)
    for(i in 2:(m+1)){
      for(j in 2:(n+1)){
        if(M[i,j] > 9 & !M_flashed[i,j]){ #OCTOPUS FLASHES!
          num_flashes <- num_flashes + 1
          #add one to neighboring octopuses
          M[(i-1):(i+1),(j-1):(j+1)] <- M[(i-1):(i+1),(j-1):(j+1)] + 1
          #reset current octopus
          M_flashed[i,j] <- TRUE
        }
      }
    }
  }
  
  M[M_flashed] <- 0
  M_flashed <- matrix(FALSE, nrow = m + 2, ncol = n + 2)
}
print(paste("The total number of flashes is", num_flashes))
t2 <- proc.time()
t2 - t1
```

--- Part Two ---

It seems like the individual flashes aren't bright enough to navigate. However, you might have a better option: the flashes seem to be synchronizing!

In the example above, the first time all octopuses flash simultaneously is step 195:

After step 193:
5877777777
8877777777
7777777777
7777777777
7777777777
7777777777
7777777777
7777777777
7777777777
7777777777

After step 194:
6988888888
9988888888
8888888888
8888888888
8888888888
8888888888
8888888888
8888888888
8888888888
8888888888

After step 195:
0000000000
0000000000
0000000000
0000000000
0000000000
0000000000
0000000000
0000000000
0000000000
0000000000

If you can calculate the exact moments when the octopuses will all flash simultaneously, you should be able to navigate through the cavern. What is the first step during which all octopuses flash?

```{r}
t1 <- proc.time()
puzzle_input <- c(4438624262, 6263251864, 2618812434, 2134264565, 1815131247, 2612457325,
8585767584, 7217134556, 2825456563, 8248473584)
#example
# puzzle_input <- c(5483143223, 2745854711, 5264556173, 6141336146, 6357385478, 4167524645, 2176841721, 6882881134, 4846848554, 5283751526)
m <- length(puzzle_input) #number of rows
n <- length(str_split(as.character(puzzle_input[1]), "")[[1]]) #number of columns

# create a matrix (but with a buffer on all 4 sides)
M <- matrix(0, nrow = m + 2, ncol = n + 2)
for(row in 2:(m+1)){
  M[row,2:(n+1)] <- as.integer(str_split(as.character(puzzle_input[row-1]), "")[[1]])
}

# simulate the octupuses flashes

M_flashed <- matrix(FALSE, nrow = m + 2, ncol = n + 2) #keep track of which octopuses flashed
for(timestep in 1:1000){ 
  
  #add one energy unit each octopus (avoid buffer in first and last rows and columns)
  M[2:(m+1),2:(n+1)] <- M[2:(m+1),2:(n+1)] + 1
  
  for(k in 1:(m*n)){ #avoid while loop
    #traverse matrix (avoid buffer in first and last rows and columns)
    for(i in 2:(m+1)){
      for(j in 2:(n+1)){
        if(M[i,j] > 9 & !M_flashed[i,j]){ #OCTOPUS FLASHES!
          #add one to neighboring octopuses
          M[(i-1):(i+1),(j-1):(j+1)] <- M[(i-1):(i+1),(j-1):(j+1)] + 1
          #reset current octopus
          M_flashed[i,j] <- TRUE
        }
      }
    }
  }
  
  #if every octopus flashed, stop here
  if(sum(M_flashed) == (m*n)){
    break
  }
  
  M[M_flashed] <- 0
  M_flashed <- matrix(FALSE, nrow = m + 2, ncol = n + 2)
}
print(paste("Every octopus flashed at time t =", timestep))
t2 <- proc.time()
t2 - t1
```

