---
title: "Day 7"
author: "Derek Sollberger"
date: "11/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
```

--- Day 7: Some Assembly Required ---

This year, Santa brought little Bobby Tables a set of wires and bitwise logic gates! Unfortunately, little Bobby is a little under the recommended age range, and he needs help assembling the circuit.

Each wire has an identifier (some lowercase letters) and can carry a 16-bit signal (a number from 0 to 65535). A signal is provided to each wire by a gate, another wire, or some specific value. Each wire can only get a signal from one source, but can provide its signal to multiple destinations. A gate provides no signal until all of its inputs have a signal.

The included instructions booklet describes how to connect the parts together: x AND y -> z means to connect wires x and y to an AND gate, and then connect its output to wire z.

For example:

    123 -> x means that the signal 123 is provided to wire x.
    x AND y -> z means that the bitwise AND of wire x and wire y is provided to wire z.
    p LSHIFT 2 -> q means that the value from wire p is left-shifted by 2 and then provided to wire q.
    NOT e -> f means that the bitwise complement of the value from wire e is provided to wire f.

Other possible gates include OR (bitwise OR) and RSHIFT (right-shift). If, for some reason, you'd like to emulate the circuit instead, almost all programming languages (for example, C, JavaScript, or Python) provide operators for these gates.

For example, here is a simple circuit:

123 -> x
456 -> y
x AND y -> d
x OR y -> e
x LSHIFT 2 -> f
y RSHIFT 2 -> g
NOT x -> h
NOT y -> i

After it is run, these are the signals on the wires:

d: 72
e: 507
f: 492
g: 114
h: 65412
i: 65079
x: 123
y: 456

In little Bobby's kit's instructions booklet (provided as your puzzle input), what signal is ultimately provided to wire a?

```{r}
full_circuit <- readLines("20151207input.txt")
example_circuit <- readLines("20151207example.txt")

bitwise_NOT <- function(x){
  # INPUT: integer (in base 10)
  # OUTPUT: integer (in base 10)
  # This function will compute the bitwise NOT operation, 
  # but for 16-bit numbers
  strtoi(paste0(rev(as.character(as.integer(!as.integer(intToBits(x)))[1:16])), collapse = ""), base = 2)
  
  # intToBits(x) %>%
  #   as.integer() %>%
  #   !(x)%>%
  #   ()[1:16] %>%
  #   as.integer() %>%
  #   as.character() %>%
  #   rev() %>%
  #   paste0(collapse = "") %>%
  #   strtoi(base = 2)
}
```

```{r}
# trying example with the bit-wise operations in R
x <- 123
y <- 456
d <- bitwAnd(x,y)
e <- bitwOr(x,y)
f <- bitwShiftL(x,2)
g <- bitwShiftR(y,2)
h <- bitwise_NOT(x)
i <- bitwise_NOT(y)
```


```{r}
N <- length(example_circuit)

# set up data frame
node_dest <- rep(NA, N)
operation <- rep(NA, N)
node1 <- rep(NA, N)
node2 <- rep(NA, N)
num_input <- rep(NA, N)


rewritten <- data.frame(node_dest, operation, node1, node2, num_input)

output_instructions <- "" #initialization

for(k in 1:N){
  split_formula <- str_split(example_circuit[k], " -> ")[[1]]
  
  RHS <- ""
  LHS <- split_formula[2]
  
  if(str_detect(split_formula[1], "AND")){
    these_instructions <- str_split(split_formula[1], " ")[[1]]
    RHS <- paste0("bitwAnd(",
                  these_instructions[1],
                  ",",
                  these_instructions[3],
                  ")")
  }
  if(str_detect(split_formula[1], "OR")){
    these_instructions <- str_split(split_formula[1], " ")[[1]]
    RHS <- paste0("bitwOr(",
                  these_instructions[1],
                  ",",
                  these_instructions[3],
                  ")")
  }
  if(str_detect(split_formula[1], "LSHIFT")){
    these_instructions <- str_split(split_formula[1], " ")[[1]]
    RHS <- paste0("bitwShiftL(",
                  these_instructions[1],
                  ",",
                  these_instructions[3],
                  ")")
  }
  if(str_detect(split_formula[1], "RSHIFT")){
    these_instructions <- str_split(split_formula[1], " ")[[1]]
    RHS <- paste0("bitwShiftR(",
                  these_instructions[1],
                  ",",
                  these_instructions[3],
                  ")")
  }
  if(str_detect(split_formula[1], "NOT")){
    these_instructions <- str_split(split_formula[1], " ")[[1]]
    RHS <- paste0("bitwise_NOT(",
                  these_instructions[2],
                  ")")
  }
  if(RHS == ""){
    RHS <- split_formula[1]
  }
  
  rewritten_instructions <- paste(LHS, "<-", RHS)
  output_instructions <- c(output_instructions, rewritten_instructions)
}

#print
output_instructions
```

```{r}
N <- length(full_circuit)

# set up data frame
node_dest <- rep(NA, N)
instructions <- rep(NA, N)
rewritten_df <- data.frame(node_dest, instructions)

for(k in 1:N){
  split_formula <- str_split(full_circuit[k], " -> ")[[1]]
  
  RHS <- ""
  LHS <- split_formula[2]
  
  if(str_detect(split_formula[1], "AND")){
    these_instructions <- str_split(split_formula[1], " ")[[1]]
    RHS <- paste0("bitwAnd(",
                  these_instructions[1],
                  ",",
                  these_instructions[3],
                  ")")
  }
  if(str_detect(split_formula[1], "OR")){
    these_instructions <- str_split(split_formula[1], " ")[[1]]
    RHS <- paste0("bitwOr(",
                  these_instructions[1],
                  ",",
                  these_instructions[3],
                  ")")
  }
  if(str_detect(split_formula[1], "LSHIFT")){
    these_instructions <- str_split(split_formula[1], " ")[[1]]
    RHS <- paste0("bitwShiftL(",
                  these_instructions[1],
                  ",",
                  these_instructions[3],
                  ")")
  }
  if(str_detect(split_formula[1], "RSHIFT")){
    these_instructions <- str_split(split_formula[1], " ")[[1]]
    RHS <- paste0("bitwShiftR(",
                  these_instructions[1],
                  ",",
                  these_instructions[3],
                  ")")
  }
  if(str_detect(split_formula[1], "NOT")){
    these_instructions <- str_split(split_formula[1], " ")[[1]]
    RHS <- paste0("bitwise_NOT(",
                  these_instructions[2],
                  ")")
  }
  if(RHS == ""){
    RHS <- split_formula[1]
  }
  
  rewritten_instructions <- paste(LHS, "<-", RHS)
  rewritten_df[k,1] <- split_formula[2]
  rewritten_df[k,2] <- rewritten_instructions
}
```

```{r}
one_letter_nodes <- rewritten_df %>%
  filter(str_length(node_dest) == 1) %>%
  arrange(node_dest)
two_letter_nodes <- rewritten_df %>%
  filter(str_length(node_dest) == 2) %>%
  arrange(node_dest)

ordered_df <- rbind(one_letter_nodes, two_letter_nodes)

writeLines(ordered_df[,2], "20151207output.txt", sep = "\n")
```

```{r}
# needed to move "a <- lx" to the end
# needed to avoid reserved words "if" and "in"

b <- 44430
c <- 0
d <- bitwShiftR(b,2)
e <- bitwShiftR(b,3)
f <- bitwShiftR(b,5)
g <- bitwOr(e,f)
h <- bitwAnd(e,f)
i <- bitwise_NOT(h)
j <- bitwAnd(g,i)
k <- bitwOr(d,j)
l <- bitwAnd(d,j)
m <- bitwise_NOT(l)
n <- bitwAnd(k,m)
o <- bitwOr(b,n)
p <- bitwAnd(b,n)
q <- bitwise_NOT(p)
r <- bitwAnd(o,q)
s <- bitwAnd(1,r)
t <- bitwShiftL(c,1)
u <- bitwOr(t,s)
v <- bitwShiftR(b,1)
w <- bitwShiftL(s,15)
x <- bitwOr(v,w)
y <- bitwShiftR(x,2)
z <- bitwShiftR(x,3)
aa <- bitwShiftR(x,5)
ab <- bitwOr(z,aa)
ac <- bitwAnd(z,aa)
ad <- bitwise_NOT(ac)
ae <- bitwAnd(ab,ad)
af <- bitwOr(y,ae)
ag <- bitwAnd(y,ae)
ah <- bitwise_NOT(ag)
ai <- bitwAnd(af,ah)
aj <- bitwOr(x,ai)
ak <- bitwAnd(x,ai)
al <- bitwise_NOT(ak)
am <- bitwAnd(aj,al)
an <- bitwAnd(1,am)
ao <- bitwShiftL(u,1)
ap <- bitwOr(ao,an)
aq <- bitwShiftR(x,1)
ar <- bitwShiftL(an,15)
as <- bitwOr(aq,ar)
at <- bitwShiftR(as,2)
au <- bitwShiftR(as,3)
av <- bitwShiftR(as,5)
aw <- bitwOr(au,av)
ax <- bitwAnd(au,av)
ay <- bitwise_NOT(ax)
az <- bitwAnd(aw,ay)
ba <- bitwOr(at,az)
bb <- bitwAnd(at,az)
bc <- bitwise_NOT(bb)
bd <- bitwAnd(ba,bc)
be <- bitwOr(as,bd)
bf <- bitwAnd(as,bd)
bg <- bitwise_NOT(bf)
bh <- bitwAnd(be,bg)
bi <- bitwAnd(1,bh)
bj <- bitwShiftL(ap,1)
bk <- bitwOr(bj,bi)
bl <- bitwShiftR(as,1)
bm <- bitwShiftL(bi,15)
bn <- bitwOr(bl,bm)
bo <- bitwShiftR(bn,2)
bp <- bitwShiftR(bn,3)
bq <- bitwShiftR(bn,5)
br <- bitwOr(bp,bq)
bs <- bitwAnd(bp,bq)
bt <- bitwise_NOT(bs)
bu <- bitwAnd(br,bt)
bv <- bitwOr(bo,bu)
bw <- bitwAnd(bo,bu)
bx <- bitwise_NOT(bw)
by <- bitwAnd(bv,bx)
bz <- bitwOr(bn,by)
ca <- bitwAnd(bn,by)
cb <- bitwise_NOT(ca)
cc <- bitwAnd(bz,cb)
cd <- bitwAnd(1,cc)
ce <- bitwShiftL(bk,1)
cf <- bitwOr(ce,cd)
cg <- bitwShiftR(bn,1)
ch <- bitwShiftL(cd,15)
ci <- bitwOr(cg,ch)
cj <- bitwShiftR(ci,2)
ck <- bitwShiftR(ci,3)
cl <- bitwShiftR(ci,5)
cm <- bitwOr(ck,cl)
cn <- bitwAnd(ck,cl)
co <- bitwise_NOT(cn)
cp <- bitwAnd(cm,co)
cq <- bitwOr(cj,cp)
cr <- bitwAnd(cj,cp)
cs <- bitwise_NOT(cr)
ct <- bitwAnd(cq,cs)
cu <- bitwOr(ci,ct)
cv <- bitwAnd(ci,ct)
cw <- bitwise_NOT(cv)
cx <- bitwAnd(cu,cw)
cy <- bitwAnd(1,cx)
cz <- bitwShiftL(cf,1)
da <- bitwOr(cz,cy)
db <- bitwShiftR(ci,1)
dc <- bitwShiftL(cy,15)
dd <- bitwOr(db,dc)
de <- bitwShiftR(dd,2)
df <- bitwShiftR(dd,3)
dg <- bitwShiftR(dd,5)
dh <- bitwOr(df,dg)
di <- bitwAnd(df,dg)
dj <- bitwise_NOT(di)
dk <- bitwAnd(dh,dj)
dl <- bitwOr(de,dk)
dm <- bitwAnd(de,dk)
dn <- bitwise_NOT(dm)
do <- bitwAnd(dl,dn)
dp <- bitwOr(dd,do)
dq <- bitwAnd(dd,do)
dr <- bitwise_NOT(dq)
ds <- bitwAnd(dp,dr)
dt <- bitwAnd(1,ds)
du <- bitwShiftL(da,1)
dv <- bitwOr(du,dt)
dw <- bitwShiftR(dd,1)
dx <- bitwShiftL(dt,15)
dy <- bitwOr(dw,dx)
dz <- bitwShiftR(dy,2)
ea <- bitwShiftR(dy,3)
eb <- bitwShiftR(dy,5)
ec <- bitwOr(ea,eb)
ed <- bitwAnd(ea,eb)
ee <- bitwise_NOT(ed)
ef <- bitwAnd(ec,ee)
eg <- bitwOr(dz,ef)
eh <- bitwAnd(dz,ef)
ei <- bitwise_NOT(eh)
ej <- bitwAnd(eg,ei)
ek <- bitwOr(dy,ej)
el <- bitwAnd(dy,ej)
em <- bitwise_NOT(el)
en <- bitwAnd(ek,em)
eo <- bitwAnd(1,en)
ep <- bitwShiftL(dv,1)
eq <- bitwOr(ep,eo)
er <- bitwShiftR(dy,1)
es <- bitwShiftL(eo,15)
et <- bitwOr(er,es)
eu <- bitwShiftR(et,2)
ev <- bitwShiftR(et,3)
ew <- bitwShiftR(et,5)
ex <- bitwOr(ev,ew)
ey <- bitwAnd(ev,ew)
ez <- bitwise_NOT(ey)
fa <- bitwAnd(ex,ez)
fb <- bitwOr(eu,fa)
fc <- bitwAnd(eu,fa)
fd <- bitwise_NOT(fc)
fe <- bitwAnd(fb,fd)
ff <- bitwOr(et,fe)
fg <- bitwAnd(et,fe)
fh <- bitwise_NOT(fg)
fi <- bitwAnd(ff,fh)
fj <- bitwAnd(1,fi)
fk <- bitwShiftL(eq,1)
fl <- bitwOr(fk,fj)
fm <- bitwShiftR(et,1)
fn <- bitwShiftL(fj,15)
fo <- bitwOr(fm,fn)
fp <- bitwShiftR(fo,2)
fq <- bitwShiftR(fo,3)
fr <- bitwShiftR(fo,5)
fs <- bitwOr(fq,fr)
ft <- bitwAnd(fq,fr)
fu <- bitwise_NOT(ft)
fv <- bitwAnd(fs,fu)
fw <- bitwOr(fp,fv)
fx <- bitwAnd(fp,fv)
fy <- bitwise_NOT(fx)
fz <- bitwAnd(fw,fy)
ga <- bitwOr(fo,fz)
gb <- bitwAnd(fo,fz)
gc <- bitwise_NOT(gb)
gd <- bitwAnd(ga,gc)
ge <- bitwAnd(1,gd)
gf <- bitwShiftL(fl,1)
gg <- bitwOr(gf,ge)
gh <- bitwShiftR(fo,1)
gi <- bitwShiftL(ge,15)
gj <- bitwOr(gh,gi)
gk <- bitwShiftR(gj,2)
gl <- bitwShiftR(gj,3)
gm <- bitwShiftR(gj,5)
gn <- bitwOr(gl,gm)
go <- bitwAnd(gl,gm)
gp <- bitwise_NOT(go)
gq <- bitwAnd(gn,gp)
gr <- bitwOr(gk,gq)
gs <- bitwAnd(gk,gq)
gt <- bitwise_NOT(gs)
gu <- bitwAnd(gr,gt)
gv <- bitwOr(gj,gu)
gw <- bitwAnd(gj,gu)
gx <- bitwise_NOT(gw)
gy <- bitwAnd(gv,gx)
gz <- bitwAnd(1,gy)
ha <- bitwShiftL(gg,1)
hb <- bitwOr(ha,gz)
hc <- bitwShiftR(gj,1)
hd <- bitwShiftL(gz,15)
he <- bitwOr(hc,hd)
hf <- bitwShiftR(he,2)
hg <- bitwShiftR(he,3)
hh <- bitwShiftR(he,5)
hi <- bitwOr(hg,hh)
hj <- bitwAnd(hg,hh)
hk <- bitwise_NOT(hj)
hl <- bitwAnd(hi,hk)
hm <- bitwOr(hf,hl)
hn <- bitwAnd(hf,hl)
ho <- bitwise_NOT(hn)
hp <- bitwAnd(hm,ho)
hq <- bitwOr(he,hp)
hr <- bitwAnd(he,hp)
hs <- bitwise_NOT(hr)
ht <- bitwAnd(hq,hs)
hu <- bitwAnd(1,ht)
hv <- bitwShiftL(hb,1)
hw <- bitwOr(hv,hu)
hx <- bitwShiftR(he,1)
hy <- bitwShiftL(hu,15)
hz <- bitwOr(hx,hy)
ia <- bitwShiftR(hz,2)
ib <- bitwShiftR(hz,3)
ic <- bitwShiftR(hz,5)
id <- bitwOr(ib,ic)
ie <- bitwAnd(ib,ic)
zz <- bitwise_NOT(ie)
ig <- bitwAnd(id,zz)
ih <- bitwOr(ia,ig)
ii <- bitwAnd(ia,ig)
ij <- bitwise_NOT(ii)
ik <- bitwAnd(ih,ij)
il <- bitwOr(hz,ik)
im <- bitwAnd(hz,ik)
yy <- bitwise_NOT(im)
io <- bitwAnd(il,yy)
ip <- bitwAnd(1,io)
iq <- bitwShiftL(hw,1)
ir <- bitwOr(iq,ip)
is <- bitwShiftR(hz,1)
it <- bitwShiftL(ip,15)
iu <- bitwOr(is,it)
iv <- bitwShiftR(iu,2)
iw <- bitwShiftR(iu,3)
ix <- bitwShiftR(iu,5)
iy <- bitwOr(iw,ix)
iz <- bitwAnd(iw,ix)
ja <- bitwise_NOT(iz)
jb <- bitwAnd(iy,ja)
jc <- bitwOr(iv,jb)
jd <- bitwAnd(iv,jb)
je <- bitwise_NOT(jd)
jf <- bitwAnd(jc,je)
jg <- bitwOr(iu,jf)
jh <- bitwAnd(iu,jf)
ji <- bitwise_NOT(jh)
jj <- bitwAnd(jg,ji)
jk <- bitwAnd(1,jj)
jl <- bitwShiftL(ir,1)
jm <- bitwOr(jl,jk)
jn <- bitwShiftR(iu,1)
jo <- bitwShiftL(jk,15)
jp <- bitwOr(jn,jo)
jq <- bitwShiftR(jp,2)
jr <- bitwShiftR(jp,3)
js <- bitwShiftR(jp,5)
jt <- bitwOr(jr,js)
ju <- bitwAnd(jr,js)
jv <- bitwise_NOT(ju)
jw <- bitwAnd(jt,jv)
jx <- bitwOr(jq,jw)
jy <- bitwAnd(jq,jw)
jz <- bitwise_NOT(jy)
ka <- bitwAnd(jx,jz)
kb <- bitwOr(jp,ka)
kc <- bitwAnd(jp,ka)
kd <- bitwise_NOT(kc)
ke <- bitwAnd(kb,kd)
kf <- bitwAnd(1,ke)
kg <- bitwShiftL(jm,1)
kh <- bitwOr(kg,kf)
ki <- bitwShiftR(jp,1)
kj <- bitwShiftL(kf,15)
kk <- bitwOr(ki,kj)
kl <- bitwShiftR(kk,2)
km <- bitwShiftR(kk,3)
kn <- bitwShiftR(kk,5)
ko <- bitwOr(km,kn)
kp <- bitwAnd(km,kn)
kq <- bitwise_NOT(kp)
kr <- bitwAnd(ko,kq)
ks <- bitwOr(kl,kr)
kt <- bitwAnd(kl,kr)
ku <- bitwise_NOT(kt)
kv <- bitwAnd(ks,ku)
kw <- bitwOr(kk,kv)
kx <- bitwAnd(kk,kv)
ky <- bitwise_NOT(kx)
kz <- bitwAnd(kw,ky)
la <- bitwAnd(1,kz)
lb <- bitwShiftL(kh,1)
lc <- bitwOr(lb,la)
ld <- bitwShiftR(kk,1)
le <- bitwShiftL(la,15)
lf <- bitwOr(ld,le)
lg <- bitwShiftR(lf,2)
lh <- bitwShiftR(lf,3)
li <- bitwShiftR(lf,5)
lj <- bitwOr(lh,li)
lk <- bitwAnd(lh,li)
ll <- bitwise_NOT(lk)
lm <- bitwAnd(lj,ll)
ln <- bitwOr(lg,lm)
lo <- bitwAnd(lg,lm)
lp <- bitwise_NOT(lo)
lq <- bitwAnd(ln,lp)
lr <- bitwOr(lf,lq)
ls <- bitwAnd(lf,lq)
lt <- bitwise_NOT(ls)
lu <- bitwAnd(lr,lt)
lv <- bitwAnd(1,lu)
lw <- bitwShiftL(lc,1)
lx <- bitwOr(lw,lv)
ly <- bitwShiftR(lf,1)
lz <- bitwShiftL(lv,15)
ma <- bitwOr(ly,lz)
a <- lx
```

```{r}
print(paste("The value of a is", a))
```

--- Part Two ---

Now, take the signal you got on wire a, override wire b to that signal, and reset the other wires (including wire a). What new signal is ultimately provided to wire a?


```{r}
b <- 3176
c <- 0
d <- bitwShiftR(b,2)
e <- bitwShiftR(b,3)
f <- bitwShiftR(b,5)
g <- bitwOr(e,f)
h <- bitwAnd(e,f)
i <- bitwise_NOT(h)
j <- bitwAnd(g,i)
k <- bitwOr(d,j)
l <- bitwAnd(d,j)
m <- bitwise_NOT(l)
n <- bitwAnd(k,m)
o <- bitwOr(b,n)
p <- bitwAnd(b,n)
q <- bitwise_NOT(p)
r <- bitwAnd(o,q)
s <- bitwAnd(1,r)
t <- bitwShiftL(c,1)
u <- bitwOr(t,s)
v <- bitwShiftR(b,1)
w <- bitwShiftL(s,15)
x <- bitwOr(v,w)
y <- bitwShiftR(x,2)
z <- bitwShiftR(x,3)
aa <- bitwShiftR(x,5)
ab <- bitwOr(z,aa)
ac <- bitwAnd(z,aa)
ad <- bitwise_NOT(ac)
ae <- bitwAnd(ab,ad)
af <- bitwOr(y,ae)
ag <- bitwAnd(y,ae)
ah <- bitwise_NOT(ag)
ai <- bitwAnd(af,ah)
aj <- bitwOr(x,ai)
ak <- bitwAnd(x,ai)
al <- bitwise_NOT(ak)
am <- bitwAnd(aj,al)
an <- bitwAnd(1,am)
ao <- bitwShiftL(u,1)
ap <- bitwOr(ao,an)
aq <- bitwShiftR(x,1)
ar <- bitwShiftL(an,15)
as <- bitwOr(aq,ar)
at <- bitwShiftR(as,2)
au <- bitwShiftR(as,3)
av <- bitwShiftR(as,5)
aw <- bitwOr(au,av)
ax <- bitwAnd(au,av)
ay <- bitwise_NOT(ax)
az <- bitwAnd(aw,ay)
ba <- bitwOr(at,az)
bb <- bitwAnd(at,az)
bc <- bitwise_NOT(bb)
bd <- bitwAnd(ba,bc)
be <- bitwOr(as,bd)
bf <- bitwAnd(as,bd)
bg <- bitwise_NOT(bf)
bh <- bitwAnd(be,bg)
bi <- bitwAnd(1,bh)
bj <- bitwShiftL(ap,1)
bk <- bitwOr(bj,bi)
bl <- bitwShiftR(as,1)
bm <- bitwShiftL(bi,15)
bn <- bitwOr(bl,bm)
bo <- bitwShiftR(bn,2)
bp <- bitwShiftR(bn,3)
bq <- bitwShiftR(bn,5)
br <- bitwOr(bp,bq)
bs <- bitwAnd(bp,bq)
bt <- bitwise_NOT(bs)
bu <- bitwAnd(br,bt)
bv <- bitwOr(bo,bu)
bw <- bitwAnd(bo,bu)
bx <- bitwise_NOT(bw)
by <- bitwAnd(bv,bx)
bz <- bitwOr(bn,by)
ca <- bitwAnd(bn,by)
cb <- bitwise_NOT(ca)
cc <- bitwAnd(bz,cb)
cd <- bitwAnd(1,cc)
ce <- bitwShiftL(bk,1)
cf <- bitwOr(ce,cd)
cg <- bitwShiftR(bn,1)
ch <- bitwShiftL(cd,15)
ci <- bitwOr(cg,ch)
cj <- bitwShiftR(ci,2)
ck <- bitwShiftR(ci,3)
cl <- bitwShiftR(ci,5)
cm <- bitwOr(ck,cl)
cn <- bitwAnd(ck,cl)
co <- bitwise_NOT(cn)
cp <- bitwAnd(cm,co)
cq <- bitwOr(cj,cp)
cr <- bitwAnd(cj,cp)
cs <- bitwise_NOT(cr)
ct <- bitwAnd(cq,cs)
cu <- bitwOr(ci,ct)
cv <- bitwAnd(ci,ct)
cw <- bitwise_NOT(cv)
cx <- bitwAnd(cu,cw)
cy <- bitwAnd(1,cx)
cz <- bitwShiftL(cf,1)
da <- bitwOr(cz,cy)
db <- bitwShiftR(ci,1)
dc <- bitwShiftL(cy,15)
dd <- bitwOr(db,dc)
de <- bitwShiftR(dd,2)
df <- bitwShiftR(dd,3)
dg <- bitwShiftR(dd,5)
dh <- bitwOr(df,dg)
di <- bitwAnd(df,dg)
dj <- bitwise_NOT(di)
dk <- bitwAnd(dh,dj)
dl <- bitwOr(de,dk)
dm <- bitwAnd(de,dk)
dn <- bitwise_NOT(dm)
do <- bitwAnd(dl,dn)
dp <- bitwOr(dd,do)
dq <- bitwAnd(dd,do)
dr <- bitwise_NOT(dq)
ds <- bitwAnd(dp,dr)
dt <- bitwAnd(1,ds)
du <- bitwShiftL(da,1)
dv <- bitwOr(du,dt)
dw <- bitwShiftR(dd,1)
dx <- bitwShiftL(dt,15)
dy <- bitwOr(dw,dx)
dz <- bitwShiftR(dy,2)
ea <- bitwShiftR(dy,3)
eb <- bitwShiftR(dy,5)
ec <- bitwOr(ea,eb)
ed <- bitwAnd(ea,eb)
ee <- bitwise_NOT(ed)
ef <- bitwAnd(ec,ee)
eg <- bitwOr(dz,ef)
eh <- bitwAnd(dz,ef)
ei <- bitwise_NOT(eh)
ej <- bitwAnd(eg,ei)
ek <- bitwOr(dy,ej)
el <- bitwAnd(dy,ej)
em <- bitwise_NOT(el)
en <- bitwAnd(ek,em)
eo <- bitwAnd(1,en)
ep <- bitwShiftL(dv,1)
eq <- bitwOr(ep,eo)
er <- bitwShiftR(dy,1)
es <- bitwShiftL(eo,15)
et <- bitwOr(er,es)
eu <- bitwShiftR(et,2)
ev <- bitwShiftR(et,3)
ew <- bitwShiftR(et,5)
ex <- bitwOr(ev,ew)
ey <- bitwAnd(ev,ew)
ez <- bitwise_NOT(ey)
fa <- bitwAnd(ex,ez)
fb <- bitwOr(eu,fa)
fc <- bitwAnd(eu,fa)
fd <- bitwise_NOT(fc)
fe <- bitwAnd(fb,fd)
ff <- bitwOr(et,fe)
fg <- bitwAnd(et,fe)
fh <- bitwise_NOT(fg)
fi <- bitwAnd(ff,fh)
fj <- bitwAnd(1,fi)
fk <- bitwShiftL(eq,1)
fl <- bitwOr(fk,fj)
fm <- bitwShiftR(et,1)
fn <- bitwShiftL(fj,15)
fo <- bitwOr(fm,fn)
fp <- bitwShiftR(fo,2)
fq <- bitwShiftR(fo,3)
fr <- bitwShiftR(fo,5)
fs <- bitwOr(fq,fr)
ft <- bitwAnd(fq,fr)
fu <- bitwise_NOT(ft)
fv <- bitwAnd(fs,fu)
fw <- bitwOr(fp,fv)
fx <- bitwAnd(fp,fv)
fy <- bitwise_NOT(fx)
fz <- bitwAnd(fw,fy)
ga <- bitwOr(fo,fz)
gb <- bitwAnd(fo,fz)
gc <- bitwise_NOT(gb)
gd <- bitwAnd(ga,gc)
ge <- bitwAnd(1,gd)
gf <- bitwShiftL(fl,1)
gg <- bitwOr(gf,ge)
gh <- bitwShiftR(fo,1)
gi <- bitwShiftL(ge,15)
gj <- bitwOr(gh,gi)
gk <- bitwShiftR(gj,2)
gl <- bitwShiftR(gj,3)
gm <- bitwShiftR(gj,5)
gn <- bitwOr(gl,gm)
go <- bitwAnd(gl,gm)
gp <- bitwise_NOT(go)
gq <- bitwAnd(gn,gp)
gr <- bitwOr(gk,gq)
gs <- bitwAnd(gk,gq)
gt <- bitwise_NOT(gs)
gu <- bitwAnd(gr,gt)
gv <- bitwOr(gj,gu)
gw <- bitwAnd(gj,gu)
gx <- bitwise_NOT(gw)
gy <- bitwAnd(gv,gx)
gz <- bitwAnd(1,gy)
ha <- bitwShiftL(gg,1)
hb <- bitwOr(ha,gz)
hc <- bitwShiftR(gj,1)
hd <- bitwShiftL(gz,15)
he <- bitwOr(hc,hd)
hf <- bitwShiftR(he,2)
hg <- bitwShiftR(he,3)
hh <- bitwShiftR(he,5)
hi <- bitwOr(hg,hh)
hj <- bitwAnd(hg,hh)
hk <- bitwise_NOT(hj)
hl <- bitwAnd(hi,hk)
hm <- bitwOr(hf,hl)
hn <- bitwAnd(hf,hl)
ho <- bitwise_NOT(hn)
hp <- bitwAnd(hm,ho)
hq <- bitwOr(he,hp)
hr <- bitwAnd(he,hp)
hs <- bitwise_NOT(hr)
ht <- bitwAnd(hq,hs)
hu <- bitwAnd(1,ht)
hv <- bitwShiftL(hb,1)
hw <- bitwOr(hv,hu)
hx <- bitwShiftR(he,1)
hy <- bitwShiftL(hu,15)
hz <- bitwOr(hx,hy)
ia <- bitwShiftR(hz,2)
ib <- bitwShiftR(hz,3)
ic <- bitwShiftR(hz,5)
id <- bitwOr(ib,ic)
ie <- bitwAnd(ib,ic)
zz <- bitwise_NOT(ie)
ig <- bitwAnd(id,zz)
ih <- bitwOr(ia,ig)
ii <- bitwAnd(ia,ig)
ij <- bitwise_NOT(ii)
ik <- bitwAnd(ih,ij)
il <- bitwOr(hz,ik)
im <- bitwAnd(hz,ik)
yy <- bitwise_NOT(im)
io <- bitwAnd(il,yy)
ip <- bitwAnd(1,io)
iq <- bitwShiftL(hw,1)
ir <- bitwOr(iq,ip)
is <- bitwShiftR(hz,1)
it <- bitwShiftL(ip,15)
iu <- bitwOr(is,it)
iv <- bitwShiftR(iu,2)
iw <- bitwShiftR(iu,3)
ix <- bitwShiftR(iu,5)
iy <- bitwOr(iw,ix)
iz <- bitwAnd(iw,ix)
ja <- bitwise_NOT(iz)
jb <- bitwAnd(iy,ja)
jc <- bitwOr(iv,jb)
jd <- bitwAnd(iv,jb)
je <- bitwise_NOT(jd)
jf <- bitwAnd(jc,je)
jg <- bitwOr(iu,jf)
jh <- bitwAnd(iu,jf)
ji <- bitwise_NOT(jh)
jj <- bitwAnd(jg,ji)
jk <- bitwAnd(1,jj)
jl <- bitwShiftL(ir,1)
jm <- bitwOr(jl,jk)
jn <- bitwShiftR(iu,1)
jo <- bitwShiftL(jk,15)
jp <- bitwOr(jn,jo)
jq <- bitwShiftR(jp,2)
jr <- bitwShiftR(jp,3)
js <- bitwShiftR(jp,5)
jt <- bitwOr(jr,js)
ju <- bitwAnd(jr,js)
jv <- bitwise_NOT(ju)
jw <- bitwAnd(jt,jv)
jx <- bitwOr(jq,jw)
jy <- bitwAnd(jq,jw)
jz <- bitwise_NOT(jy)
ka <- bitwAnd(jx,jz)
kb <- bitwOr(jp,ka)
kc <- bitwAnd(jp,ka)
kd <- bitwise_NOT(kc)
ke <- bitwAnd(kb,kd)
kf <- bitwAnd(1,ke)
kg <- bitwShiftL(jm,1)
kh <- bitwOr(kg,kf)
ki <- bitwShiftR(jp,1)
kj <- bitwShiftL(kf,15)
kk <- bitwOr(ki,kj)
kl <- bitwShiftR(kk,2)
km <- bitwShiftR(kk,3)
kn <- bitwShiftR(kk,5)
ko <- bitwOr(km,kn)
kp <- bitwAnd(km,kn)
kq <- bitwise_NOT(kp)
kr <- bitwAnd(ko,kq)
ks <- bitwOr(kl,kr)
kt <- bitwAnd(kl,kr)
ku <- bitwise_NOT(kt)
kv <- bitwAnd(ks,ku)
kw <- bitwOr(kk,kv)
kx <- bitwAnd(kk,kv)
ky <- bitwise_NOT(kx)
kz <- bitwAnd(kw,ky)
la <- bitwAnd(1,kz)
lb <- bitwShiftL(kh,1)
lc <- bitwOr(lb,la)
ld <- bitwShiftR(kk,1)
le <- bitwShiftL(la,15)
lf <- bitwOr(ld,le)
lg <- bitwShiftR(lf,2)
lh <- bitwShiftR(lf,3)
li <- bitwShiftR(lf,5)
lj <- bitwOr(lh,li)
lk <- bitwAnd(lh,li)
ll <- bitwise_NOT(lk)
lm <- bitwAnd(lj,ll)
ln <- bitwOr(lg,lm)
lo <- bitwAnd(lg,lm)
lp <- bitwise_NOT(lo)
lq <- bitwAnd(ln,lp)
lr <- bitwOr(lf,lq)
ls <- bitwAnd(lf,lq)
lt <- bitwise_NOT(ls)
lu <- bitwAnd(lr,lt)
lv <- bitwAnd(1,lu)
lw <- bitwShiftL(lc,1)
lx <- bitwOr(lw,lv)
ly <- bitwShiftR(lf,1)
lz <- bitwShiftL(lv,15)
ma <- bitwOr(ly,lz)
a <- lx
```

```{r}
print(paste("The value of a is", a))
```