---
title: "Day 16"
author: "Derek Sollberger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
```

--- Day 16: Permutation Promenade ---

You come upon a very unusual sight; a group of programs here appear to be dancing.

There are sixteen programs in total, named a through p. They start by standing in a line: a stands in position 0, b stands in position 1, and so on until p, which stands in position 15.

The programs' dance consists of a sequence of dance moves:

    Spin, written sX, makes X programs move from the end to the front, but maintain their order otherwise. (For example, s3 on abcde produces cdeab).
    Exchange, written xA/B, makes the programs at positions A and B swap places.
    Partner, written pA/B, makes the programs named A and B swap places.

For example, with only five programs standing in a line (abcde), they could do the following dance:

    s1, a spin of size 1: eabcd.
    x3/4, swapping the last two programs: eabdc.
    pe/b, swapping programs e and b: baedc.

After finishing their dance, the programs end up in order baedc.

You watch the dance for a while and record their dance moves (your puzzle input). In what order are the programs standing after their dance?

```{r}
puzzle_input <- readLines("20171216input.txt")
```
```{r}
# example
# dancers <- letters[1:5]
# D <- length(dancers)
# example_input <- "s1,x3/4,pe/b"
# instructions <- str_split(example_input, ",")[[1]]
# N <- length(instructions)

# main puzzle
t1 <- proc.time()
dancers <- letters[1:16]
D <- length(dancers)
instructions <- str_split(puzzle_input, ",")[[1]]
N <- length(instructions)

for(i in 1:N){
  # spin
  if(str_starts(instructions[i], "s")){
    this_offset <- as.numeric(str_extract(instructions[i], "(\\d)+")) %% D
    if(this_offset == 0){ this_offset <- D }
    dancers <- c(tail(dancers, this_offset), head(dancers, -this_offset))
  }
  
  # exchange
  if(str_starts(instructions[i], "x")){
    this_pair <- as.numeric(
      str_extract_all(instructions[i], "(\\d)+")[[1]]) + 1
    new_dancers <- dancers
    new_dancers[this_pair[1]] <- dancers[this_pair[2]]
    new_dancers[this_pair[2]] <- dancers[this_pair[1]]
    dancers <- new_dancers
  }
  
  # partners
  if(str_starts(instructions[i], "p")){
    this_split <- str_split(instructions[i], "/")[[1]]
    partner1 <- str_sub(this_split[1], 2)
    partner2 <- this_split[2]
    place1 <- which.max(dancers == partner1)
    place2 <- which.max(dancers == partner2)
    new_dancers <- dancers
    new_dancers[place1] <- dancers[place2]
    new_dancers[place2] <- dancers[place1]
    dancers <- new_dancers
  }
}
print("The dancers are now in order")
print(paste0(dancers, collapse = ""))
t2 <- proc.time()
t2 - t1
# kpbodeajhlicngmf was the correct answer
```

--- Part Two ---

Now that you're starting to get a feel for the dance moves, you turn your attention to the dance as a whole.

Keeping the positions they ended up in from their previous dance, the programs perform it again and again: including the first dance, a total of one billion (1000000000) times.

In the example above, their second dance would begin with the order baedc, and use the same dance moves:

    s1, a spin of size 1: cbaed.
    x3/4, swapping the last two programs: cbade.
    pe/b, swapping programs e and b: ceadb.

In what order are the programs standing after their billion dances?

```{r}
# main puzzle
t1 <- proc.time()
dancers <- letters[1:16]
D <- length(dancers)
instructions <- str_split(puzzle_input, ",")[[1]]
N <- length(instructions)

for(k in 1:6){
  new_dancers <- rep(NA, D)
  new_dancers <- dancers[c(11,16,2,15,4,5,1,10,8,12,9,3,14,7,13,6)]
  # if(identical(new_dancers, letters[1:16])){
  #   print(paste("Original position reached at iteration", k))
  #   break
  # }
  # cycle was found after 16 iterations, so change max to
  # 1e9 %% 17?
  dancers <- new_dancers
}

print("The dancers are now in order")
print(paste0(dancers, collapse = ""))
t2 <- proc.time()
t2 - t1
```

