---
title: "Day 7"
author: "Derek Sollberger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
```

--- Day 7: Recursive Circus ---

Wandering further through the circuits of the computer, you come upon a tower of programs that have gotten themselves into a bit of trouble. A recursive algorithm has gotten out of hand, and now they're balanced precariously in a large tower.

One program at the bottom supports the entire tower. It's holding a large disc, and on the disc are balanced several more sub-towers. At the bottom of these sub-towers, standing on the bottom disc, are other programs, each holding their own disc, and so on. At the very tops of these sub-sub-sub-...-towers, many programs stand simply keeping the disc below them balanced but with no disc of their own.

You offer to help, but first you need to understand the structure of these towers. You ask each program to yell out their name, their weight, and (if they're holding a disc) the names of the programs immediately above them balancing on that disc. You write this information down (your puzzle input). Unfortunately, in their panic, they don't do this in an orderly fashion; by the time you're done, you're not sure which program gave which information.

For example, if your list is the following:

pbga (66)
xhth (57)
ebii (61)
havc (66)
ktlj (57)
fwft (72) -> ktlj, cntj, xhth
qoyq (66)
padx (45) -> pbga, havc, qoyq
tknk (41) -> ugml, padx, fwft
jptl (61)
ugml (68) -> gyxo, ebii, jptl
gyxo (61)
cntj (57)

...then you would be able to recreate the structure of the towers that looks like this:

                gyxo
              /     
         ugml - ebii
       /      \     
      |         jptl
      |        
      |         pbga
     /        /
tknk --- padx - havc
     \        \
      |         qoyq
      |             
      |         ktlj
       \      /     
         fwft - cntj
              \     
                xhth

In this example, tknk is at the bottom of the tower (the bottom program), and is holding up ugml, padx, and fwft. Those programs are, in turn, holding up other programs; in this example, none of those programs are holding up any other programs, and are all the tops of their own towers. (The actual tower balancing in front of you is much larger.)

Before you're ready to help them, you need to make sure your information is correct. What is the name of the bottom program?

```
I just searched the program names that appeared before "->" until I found one that only appears once: eqgvf
```

--- Part Two ---

The programs explain the situation: they can't get down. Rather, they could get down, if they weren't expending all of their energy trying to keep the tower balanced. Apparently, one program has the wrong weight, and until it's fixed, they're stuck here.

For any program holding a disc, each program standing on that disc forms a sub-tower. Each of those sub-towers are supposed to be the same weight, or the disc itself isn't balanced. The weight of a tower is the sum of the weights of the programs in that tower.

In the example above, this means that for ugml's disc to be balanced, gyxo, ebii, and jptl must all have the same weight, and they do: 61.

However, for tknk to be balanced, each of the programs standing on its disc and all programs above it must each match. This means that the following sums must all be the same:

    ugml + (gyxo + ebii + jptl) = 68 + (61 + 61 + 61) = 251
    padx + (pbga + havc + qoyq) = 45 + (66 + 66 + 66) = 243
    fwft + (ktlj + cntj + xhth) = 72 + (57 + 57 + 57) = 243

As you can see, tknk's disc is unbalanced: ugml's stack is heavier than the other two. Even though the nodes above ugml are balanced, ugml itself is too heavy: it needs to be 8 units lighter for its stack to weigh 243 and keep the towers balanced. If this change were made, its weight would be 60.

Given that exactly one program is the wrong weight, what would its weight need to be to balance the entire tower?

```{r}
puzzle_input <- readLines("20171207input.txt")
N <- length(puzzle_input)
```

```{r}
programs <- rep(NA, N)
weight_self <- rep(NA, N)
supports <- rep(NA, N)
df <- data.frame(programs, weight_self, supports)

for(i in 1:N){
  instruction_split <- str_split(puzzle_input[i], " -> ")[[1]]
  df$programs[i] <- str_split(instruction_split[1], " \\(")[[1]][1]
  df$weight_self[i] <- str_extract(instruction_split[1], "(\\d+)")
  df$supports[i] <- instruction_split[2]
}
```

```{r}
if(!is.na(df$supports[1])){
   #move row to bottom
   df <- df[c(2:N, i), ]
}

for(k in 1:100){ #run shuffling algorithm several times
  for(i in 2:(N-1)){
    if(!is.na(df$supports[i])){
      support_list <- str_split(df$supports[i], ", ")[[1]]
      n <- length(support_list)
      for(j in 1:n){
        if(!(support_list[j] %in% df$programs[1:(i-1)])){
          #move row to bottom
          df <- df[c(1:(i-1), (i+1):N, i), ]
        }
      }
    }
  }
}

#crude fix for last row
df <- df[c(1:(N-3), N, (N-2):(N-1)), ]
```

```{r}
new_instructions <- ""
for(i in 1:N){
  add_supports <- ""
  if(!is.na(df$supports[i])){
    add_supports <- paste("+", str_replace_all(df$supports[i], ",", " + "),
                          "\n", "var(c(", df$supports, "))")
  }
  this_line <- paste(df$programs[i], "<-", df$weight_self[i], add_supports)
  new_instructions <- c(new_instructions, this_line)
}
write_lines(new_instructions, "20171207_new_instructions.txt")

# missed above, so added
# bqyqwn <- 68 + wscqe + cwxspl + syogw + xnxudsh
```

```{r}

ddswb <- 34 
hnkvw <- 320 
ojzogs <- 79 
zjpkoar <- 40 
ilsgqdw <- 37 
bmlfd <- 43 
nxqwag <- 81 
ubbwvvl <- 44 
uunpyb <- 82 
dmidur <- 57 
fklfgd <- 65 
vcjfe <- 22 
rkjfx <- 33 
ioglbe <- 88 
tymkixg <- 36 
qbhhtw <- 90 
afbrhy <- 79 
nswcez <- 47 
wzjonl <- 16 
xviqup <- 95 
vtokttx <- 55 
boalcdm <- 13 
yseezd <- 11 
qcmwsvm <- 37 
sjoov <- 6 
fosdh <- 90 
pjkyydd <- 15 
wvmtyez <- 46 
urwaex <- 59 
iktget <- 33 
gxtfu <- 57 
rkctaq <- 20 
sbwoye <- 50 
tduix <- 18 
ofbqk <- 148 
ewluica <- 22 
uezwz <- 20 
ujgeml <- 94 
jkoir <- 6 
kemkz <- 98 
dtqgdc <- 90 
umrqldb <- 99 
oyurlfq <- 57 
nlajc <- 85 
xnszvs <- 65 
fmdba <- 44 
zgxbjj <- 96 
jncjxue <- 66 
iftcnc <- 50 
dqdmie <- 82 
jlsmm <- 32 
vsjjq <- 30 
yjnzgl <- 67 
ooczbeh <- 36 
ndmkbul <- 55 
wktzh <- 93 
ystgyng <- 50 
jclntv <- 87 
acjedny <- 49 
mfwuqb <- 60 
ayuma <- 81 
qqmchoq <- 7 
srypr <- 64 
vkihn <- 36 
symdxo <- 8 
nduildw <- 30 
nxhfgx <- 18 
chgfb <- 16 
gugaaiq <- 40 
euvrdg <- 79 
ckzaj <- 67 
jupzgjm <- 90 
llalaql <- 27 
qdnghwn <- 16 
enmjipp <- 95 
jtbrjm <- 18 
rlxto <- 83 
xpotznx <- 41 
lslhph <- 59 
vwkegak <- 92 
bsvgagj <- 41 
xfbfvuu <- 65 
gpwbcg <- 5 
elysbb <- 81 
engpslr <- 85 
tqbtvqc <- 31 
faqngn <- 98 
spnllx <- 76 
mbbuqoi <- 35 
zjwqc <- 26 
yjwvnhl <- 219 
fykgs <- 43 
rdzsisf <- 59 
rnkbli <- 84 
gsyeaes <- 26 
xulgx <- 5 
wleavrv <- 16 
volompq <- 96 
qmshniw <- 443 
lrmnb <- 98 
yijwaf <- 32 
paoxff <- 98 
ulbute <- 95 
ynlzggk <- 36 
pggjo <- 19 
jajdx <- 19 
ackyirz <- 16 
gtiopzn <- 49 
idvnxqp <- 5 
pmdcajx <- 7 
nzfko <- 7 
vcnael <- 95 
nxuer <- 13 
aqxaecv <- 297 
ppjapn <- 27 
jiwfia <- 13 
jthpj <- 44 
dtijfqc <- 76 
vgfhpwb <- 99 
qtzelhq <- 62 
resoi <- 25 
hgwkyu <- 91 
hsfhqhe <- 17 
metif <- 68 
rqaxn <- 96 
vxnsova <- 5 
xsxnizc <- 50 
hskhxx <- 36 
ogcmm <- 57 
drnki <- 25 
ecjxs <- 197 
dnniv <- 16 
hpwaux <- 97 
sadyx <- 59 
yytmc <- 66 
hmbuu <- 16 
ezzpr <- 94 
earmfdp <- 83 
mxpgmte <- 5 
dsezlbe <- 54 
msqxo <- 47 
jacmj <- 29 
hjsrvj <- 64 
prcxksh <- 20 
pkipy <- 95 
czbydf <- 75 
bpguivk <- 32 
utapoe <- 84 
qjdou <- 87 
njhlrtm <- 38 
pkzwosb <- 36 
wpphwjv <- 83 
teqej <- 66 
lneyu <- 64 
dicalia <- 44 
opzxu <- 46 
gasejp <- 15 
llikz <- 22 
uhtnl <- 5 
kntwy <- 25 
uhnlse <- 57 
yaxeloo <- 8 
tzbmy <- 71 
pjecv <- 56 
gsaqlad <- 70 
gmzcjo <- 76 
twyhx <- 86 
wltax <- 21 
xfyqx <- 20 
tnrrui <- 57 
gmveehb <- 46 
yxxgd <- 296 
utiwem <- 81 
jhydv <- 95 
kwhgtnk <- 78 
npxexv <- 66 
lycfbo <- 89 
vezhmd <- 47 
upurae <- 33 + rdzsisf +  lslhph +  urwaex 
 var(c( rdzsisf, lslhph, urwaex ))
uvkwg <- 88 
wapknj <- 2443 + xulgx +  idvnxqp 
 var(c( xulgx, idvnxqp ))
xlacmu <- 87 
aesiz <- 86 
jjiikvt <- 25 
prjpkm <- 22 
rqhfxty <- 53 
qhpvuzi <- 75 
cvxqnh <- 53 
lrtzi <- 53 
pjlxc <- 62 
slufzy <- 34 
zkdwz <- 20 
myhnuu <- 66 
jrsfu <- 79 
hjbfr <- 20 
lltgjo <- 32 
picnycy <- 33 
kplvk <- 24 
qcmtgvr <- 12 
gvdlvgg <- 20 
hjzgzoe <- 38 
hjqsj <- 85 
fgzgfc <- 44 
ejtdhx <- 57 
dinpz <- 20 
yxxdal <- 44 
kvylxr <- 92 
xbtjl <- 294 
dyfzj <- 95 
utsnke <- 57 
sxjlr <- 91 
xguqqao <- 98 
dhvsrt <- 7 
mjqkaqs <- 50 
dcriqom <- 79 
zppujp <- 34 
aqacmr <- 18 
rkazlp <- 38 
fysxwmd <- 36 
wqshh <- 460 
rztur <- 86 
ocdpr <- 44 
gacseeh <- 89 
xzsmw <- 24 
liujmc <- 41 
myetb <- 26 
xawzbz <- 64 
mlzmimg <- 24 
xhkwrs <- 324 
tdkdiqe <- 87 
ukgbdva <- 65 
qmcphv <- 32 
fmhfnhv <- 24 
nhjtvk <- 47 
sbmqxu <- 25 
ewobo <- 25 
hvgozl <- 17 
ysxvzdu <- 65 
sfazwz <- 68 
hlwesdc <- 24 
wgbvle <- 422 
ugzmeb <- 77 
auezw <- 95 
lkuwf <- 95 
koxzeku <- 79 
vehziv <- 67 
kjwfeoj <- 82 
lcnsf <- 22 
zzwmn <- 40 
jpmrs <- 94 
pcvfgti <- 29 
ucpkpfb <- 181 
mgcpbp <- 34 
jxyakll <- 97 
fdeowqq <- 77 
zxmwy <- 35 
yudzxoa <- 84 
wwbebbr <- 95 
wgtnepi <- 66 
wkkdc <- 40 
xlccnkv <- 13 
yxrhis <- 8 
uwjobru <- 22 
csfzl <- 64 
dspkk <- 66 
xwknab <- 28 
fmdsfus <- 98 
vpirde <- 64 
nulax <- 10 
ymkgx <- 23 
psetts <- 80 
ucfipvi <- 67 
snqnh <- 95 
gachtig <- 49 
qxnvoh <- 214 
ihtfo <- 87 
qxrspl <- 95 
nnciip <- 95 
iqmpc <- 53 
rnbvid <- 49 
xukfcd <- 28 
bjadb <- 49 
niifkl <- 90 
kxrbd <- 20 
lrcstqr <- 25 
gackyrq <- 22 
llpbev <- 29 
efqjc <- 15 
efvqwzi <- 7 
epqowm <- 45 
fjdlcbh <- 139 
csfjt <- 92 
wclnx <- 91 
aditjen <- 7 
auhnyn <- 25 
nvigovl <- 98 
yeyxf <- 16 
jaasf <- 95 
dwrtbo <- 87 
ciquv <- 24 
psjec <- 97 
xyxidwf <- 24 
esized <- 69 
towuup <- 68 
trszl <- 44 
tebkmhi <- 25 
ltytvpb <- 90 
msdcz <- 24 
csdap <- 27 
alneot <- 25 
tyvjs <- 12 
gbckkh <- 20 
dgqfrv <- 35 
oozbxqb <- 92 
amklj <- 23 
tvbgr <- 80 
qsrxouy <- 96 
rauho <- 72 
tzrqwt <- 77 
pdazlob <- 16 
uhuees <- 19 
qebqgxd <- 94 
ffhsg <- 79 
hzuyh <- 50 
oiuztx <- 92 
wjdeth <- 5 
bjops <- 18 
bwvlnlt <- 38 
baedi <- 89 
zvtfm <- 41 
aizdt <- 54 + zjpkoar +  zzwmn 
 var(c( zjpkoar, zzwmn ))
ffzzzc <- 67 
dtazpb <- 85 
oxlzlor <- 16 
bvejguc <- 68 
zpaivc <- 19 
nofpmh <- 290 + efvqwzi +  pmdcajx 
 var(c( efvqwzi, pmdcajx ))
lkdyq <- 57 
fuaio <- 16 
bgnfpcr <- 75 
jeadxth <- 22 
rteyxil <- 334 
clqfbvi <- 66 
lfdery <- 33 
ijtis <- 85 
rmshk <- 24 
kptdem <- 97 
bzqjb <- 61 
idokb <- 95 
kynezb <- 79 
athuly <- 98 
vazoq <- 248 
rxcgwd <- 38 
tfezflv <- 95 
vvzmdr <- 18 
gwuixj <- 31 
svtqdbd <- 61 
nqfqgj <- 22 + niifkl +  fosdh 
 var(c( niifkl, fosdh ))
lxzvkpk <- 458 
qqqbf <- 52 
fpeiss <- 76 
qsstg <- 57 
brvxyyg <- 30 
oqmxlu <- 39 
olebdv <- 65 
vmsoc <- 61 + euvrdg +  dcriqom 
 var(c( euvrdg, dcriqom ))
qiznqk <- 77 
kspsz <- 112 + tduix +  aqacmr 
 var(c( tduix, aqacmr ))
kybtla <- 20 
jmvcgo <- 88 
hjzkjuw <- 76 
qinne <- 35 
mrxdp <- 22 
pznkdcr <- 67 
kmvvxon <- 16 
xlaqpg <- 81 
nhtfal <- 23 
pywtg <- 16 
omslcm <- 34 
sxdngcn <- 6 
insleik <- 209 
wxvjyn <- 63 
ntezb <- 25 
klkwsa <- 57 
figvamv <- 32 
vrtgae <- 66 
okurp <- 5 
rndvcf <- 39 
jduij <- 65 
glrmew <- 67 
uqmaluk <- 70 
qstlv <- 75 + qmcphv +  bpguivk 
 var(c( qmcphv, bpguivk ))
gjgrhw <- 94 
paeel <- 87 
xiqthw <- 87 
mevujmj <- 83 
xctuhbx <- 72 
njxrb <- 60 
qotyaq <- 7 
dnbejx <- 70 
nggubzz <- 84 
uuunqk <- 9 
lxttas <- 133 
xvdmeo <- 7 
gkkit <- 84 
wexncjl <- 295 
jgsmitm <- 173 
ibqrz <- 96 
xfcktep <- 13 
eteyvz <- 23 
eohjrft <- 55 
wqilo <- 32 
tbmyue <- 54 
uuxpvxf <- 114 + nhjtvk +  vezhmd 
 var(c( nhjtvk, vezhmd ))
yrolq <- 64 
usevjww <- 15 
zxygmaq <- 10 
tdmirp <- 75 
zodwdpq <- 93 
uyhrcfl <- 76 
mqgbif <- 73 
petvm <- 24 
nxxlbnt <- 34 
kpbawoe <- 26 
recxxj <- 53 
ejhojmh <- 56 
dnovxt <- 44 
crslvc <- 69 
fnayjx <- 64 
lytrpo <- 47 
cnffplo <- 166 + fuaio +  pywtg 
 var(c( fuaio, pywtg ))
xznhz <- 54 
mclqrmw <- 15 
uqynzy <- 80 
yhukqf <- 50 
upozls <- 20 
qladcl <- 44 
wjgjhc <- 5 
fucxhz <- 93 
nayvvrh <- 25 + ejtdhx +  oyurlfq 
 var(c( ejtdhx, oyurlfq ))
wvvivob <- 6 
wksjlix <- 81 
fzyim <- 20 
zfipnh <- 21 
uopxjn <- 64 
gaxdd <- 25 
dzwfsix <- 87 
nrxyk <- 29 
vbjpz <- 13 
nfkfex <- 134 + vsjjq +  nduildw 
 var(c( vsjjq, nduildw ))
cpujsn <- 22 
weipv <- 91 
wjceo <- 184 + resoi +  gaxdd 
 var(c( resoi, gaxdd ))
lnfjj <- 50 
bmmzqxb <- 81 
jimltsw <- 16 
nptqxtu <- 95 
rpwdig <- 10 
whrinsh <- 34 
coapfqs <- 20 
souqntk <- 98 
rpgzp <- 44 
wgupfl <- 65 
jxawk <- 34 
zeolg <- 214 
qzvobt <- 53 
rkuzg <- 83 
gdvuw <- 179 + vrtgae +  teqej 
 var(c( vrtgae, teqej ))
pinqp <- 59 
xtnvc <- 99 
gppwsbp <- 35 
jmuqp <- 44 
mieuhe <- 97 
donzb <- 407 
igbzlxl <- 27 
invqsb <- 5 
xxdcgas <- 30 + auezw +  wwbebbr 
 var(c( auezw, wwbebbr ))
qhkhuow <- 44 
sgytd <- 62 
iiuqtqw <- 45 
isrgei <- 173 
sebvkw <- 214 + qotyaq +  dhvsrt 
 var(c( qotyaq, dhvsrt ))
obzrtar <- 62 
nakli <- 50 
bzmqg <- 374 
pdzooh <- 13 
scvhl <- 28 
himyt <- 41 
tptqtpq <- 57 
zkwajbd <- 61 
lpsaav <- 94 
bmcwbsi <- 96 
rtbcch <- 64 
nadbegy <- 66 
jyazs <- 74 
xcmzg <- 30 
kirnt <- 134 + bmlfd +  fykgs 
 var(c( bmlfd, fykgs ))
xhgza <- 16 
xajzwdh <- 65 
tpcbt <- 81 
zmoxc <- 99 
bcbftl <- 38 
arrvxtj <- 71 
imiimt <- 22 
tqelui <- 86 
sibvtz <- 79 
pchfpqp <- 86 
faopk <- 147 + figvamv +  jlsmm 
 var(c( figvamv, jlsmm ))
nreuwh <- 62 
eourjv <- 204 + kpbawoe +  zjwqc 
 var(c( kpbawoe, zjwqc ))
ssvfbfs <- 96 
dxlpr <- 8 
byojhfr <- 97 
nlwakg <- 161 
ztxcc <- 68 
zisdcf <- 18 
itutdol <- 55 
cseckdt <- 11 
zqwqcnq <- 37 
fhqhu <- 188 + igbzlxl +  llalaql +  ppjapn +  csdap 
 var(c( igbzlxl, llalaql, ppjapn, csdap ))
kzsoasg <- 70 
qvaayfl <- 80 
bcesxba <- 444 
hnyjyga <- 95 
idmqvse <- 32 
ozyqew <- 95 
afntec <- 96 
cgveph <- 71 
isxysg <- 8 
cqlvkj <- 65 
ykitsi <- 63 
bjhxuqc <- 53 
uahcz <- 50 
vbejzc <- 38 
dofou <- 44 
uijffti <- 65 
nkzgse <- 7 
zylotoh <- 7 
phcqntb <- 76 
amcixd <- 29 
jlsryou <- 95 
rceksnv <- 57 
tlpfggk <- 25 
fyaxzl <- 18 
qpzszp <- 65 
nevrtrs <- 83 
jyftx <- 76 
zacwh <- 5 
oqsdzpe <- 50 
lpdhdvy <- 65 
kmhug <- 64 
bvtgvu <- 94 
supjxi <- 21 
rfqddq <- 53 
ptxfk <- 73 
lofybs <- 89 
znagag <- 89 
zmciyk <- 19 + mieuhe +  psjec 
 var(c( mieuhe, psjec ))
dhobx <- 95 
tmzxl <- 23 
tmkli <- 66 
exuyuk <- 82 
xadjes <- 10 
khfzis <- 48 + ztxcc +  bvejguc 
 var(c( ztxcc, bvejguc ))
gvtcy <- 39 
qkzpcyy <- 96 
fifams <- 415 
wfgzr <- 82 
pplis <- 9 
bzfkt <- 66 + rauho +  xctuhbx 
 var(c( rauho, xctuhbx ))
stofi <- 16 
qjzss <- 31 
gdamb <- 99 
vktjns <- 45 
nyouzhg <- 17 + lofybs +  baedi 
 var(c( lofybs, baedi ))
mklqpn <- 9 
xkymbld <- 37 + mqgbif +  ptxfk 
 var(c( mqgbif, ptxfk ))
qjeuqe <- 92 
cwfydbh <- 369 + qcmtgvr +  tyvjs 
 var(c( qcmtgvr, tyvjs ))
cjjytkr <- 305 + trszl +  fgzgfc 
 var(c( trszl, fgzgfc ))
qsakd <- 82 
rrzyzq <- 76 
eaufvn <- 78 + dtazpb +  nlajc 
 var(c( dtazpb, nlajc ))
uvpuko <- 91 
mgcnohi <- 69 
klege <- 92 
mulvwn <- 290 + ubbwvvl +  jmuqp +  dnovxt 
 var(c( ubbwvvl, jmuqp, dnovxt ))
trlcj <- 245 
flgpg <- 188 + jimltsw +  xhgza 
 var(c( jimltsw, xhgza ))
wawtrk <- 80 
otlvq <- 40 + hjsrvj +  rtbcch +  vpirde 
 var(c( hjsrvj, rtbcch, vpirde ))
fkzxb <- 89 
rftuf <- 43 
uolqt <- 9 
juoaena <- 94 
mlalrpt <- 166 + jtbrjm +  fyaxzl 
 var(c( jtbrjm, fyaxzl ))
srvqt <- 155 + jyftx +  dtijfqc +  phcqntb +  uyhrcfl 
 var(c( jyftx, dtijfqc, phcqntb, uyhrcfl ))
ynvesz <- 88 + rqaxn +  afntec 
 var(c( rqaxn, afntec ))
ibbxb <- 194 + imiimt +  llikz 
 var(c( imiimt, llikz ))
ybosi <- 18 
opvrgh <- 79 
blyms <- 46 
fitha <- 68 + faqngn +  xguqqao +  athuly +  fmdsfus 
 var(c( faqngn, xguqqao, athuly, fmdsfus ))
xyang <- 225 + njxrb +  mfwuqb 
 var(c( njxrb, mfwuqb ))
yzjhmqq <- 171 + dinpz +  upozls 
 var(c( dinpz, upozls ))
qfopwo <- 79 
tvwmzal <- 236 + cpujsn +  ewluica 
 var(c( cpujsn, ewluica ))
dodrbx <- 227 + uhtnl +  vxnsova 
 var(c( uhtnl, vxnsova ))
sfdclf <- 238 
ceagctk <- 255 + bwvlnlt +  rkazlp +  vbejzc +  njhlrtm 
 var(c( bwvlnlt, rkazlp, vbejzc, njhlrtm ))
gbbpqj <- 34 
mncivb <- 88 
hbhfdlh <- 89 
chcjswk <- 89 
vrmqym <- 97 + lrmnb +  nvigovl 
 var(c( lrmnb, nvigovl ))
yapfoe <- 59 
mgbwxw <- 52 + afbrhy +  jrsfu 
 var(c( afbrhy, jrsfu ))
ysbhsn <- 181 + wleavrv +  kmvvxon 
 var(c( wleavrv, kmvvxon ))
dndrnc <- 25 
dgtxjuj <- 97 
qnhbmzp <- 118 + wktzh +  zodwdpq 
 var(c( wktzh, zodwdpq ))
rodntl <- 60 
tykcvi <- 9 
zdelqb <- 155 + jaasf +  snqnh 
 var(c( jaasf, snqnh ))
akwcv <- 65 
mabcqcb <- 194 + itutdol +  vtokttx 
 var(c( itutdol, vtokttx ))
trvtkqj <- 49 
qkjdjpe <- 44 
lvkeqm <- 192 + lneyu +  uopxjn 
 var(c( lneyu, uopxjn ))
qvrvcu <- 19 
ocgbpez <- 114 + gjgrhw +  bvtgvu 
 var(c( gjgrhw, bvtgvu ))
mehuxr <- 53 
hshafy <- 65 
pulql <- 59 + akwcv +  uijffti +  ysxvzdu +  xfbfvuu 
 var(c( akwcv, uijffti, ysxvzdu, xfbfvuu ))
hlbotz <- 13 
zhrdezs <- 28 
ikvzux <- 44 
ytsmc <- 22 
qllime <- 202 + dspkk +  wgtnepi 
 var(c( dspkk, wgtnepi ))
ksbwlir <- 36 
hzawlq <- 16 
kgwftu <- 87 
dyazidi <- 219 
urdgg <- 166 + srypr +  fnayjx 
 var(c( srypr, fnayjx ))
xboqv <- 17 
osnkrr <- 53 
aahqvqy <- 74 
dptekr <- 112 + xiqthw +  ihtfo 
 var(c( xiqthw, ihtfo ))
dsptf <- 8 
vkaxa <- 83 
kwnqw <- 49 
cuscn <- 114 + fmdba +  ikvzux 
 var(c( fmdba, ikvzux ))
egizyey <- 78 + metif +  towuup 
 var(c( metif, towuup ))
avecfoq <- 95 
kymbvpp <- 67 + qzvobt +  osnkrr 
 var(c( qzvobt, osnkrr ))
xbccghy <- 66 
cfwxyl <- 210 
vfupn <- 32 
vspqi <- 201 + wzjonl +  dnniv +  pdazlob 
 var(c( wzjonl, dnniv, pdazlob ))
amqvr <- 274 + aditjen +  xvdmeo 
 var(c( aditjen, xvdmeo ))
vbqpp <- 35 
qmifv <- 163 + hzawlq +  ackyirz 
 var(c( hzawlq, ackyirz ))
ftezz <- 24 
iuuouds <- 14 
jwjuth <- 390 + hmbuu +  oxlzlor 
 var(c( hmbuu, oxlzlor ))
stqmwk <- 96 + vcnael +  dhobx 
 var(c( vcnael, dhobx ))
qdfvw <- 69 + hgwkyu +  uvpuko +  wclnx 
 var(c( hgwkyu, uvpuko, wclnx ))
inddon <- 18 
nzhbfc <- 207 + hvgozl +  hsfhqhe 
 var(c( hvgozl, hsfhqhe ))
gqeiy <- 257 + opvrgh +  sibvtz 
 var(c( opvrgh, sibvtz ))
xgzluhk <- 21 
tqplqn <- 36 
aifpxqv <- 89 
rqyxzn <- 34 
vdcfh <- 68 
fyxttac <- 39 
hbbpp <- 95 
asevk <- 44 + hpwaux +  byojhfr 
 var(c( hpwaux, byojhfr ))
fqsjaz <- 10 
wozakri <- 251 + mgcpbp +  omslcm 
 var(c( mgcpbp, omslcm ))
dhwakp <- 26 
bixxpzd <- 87 
jllqymn <- 110 + dzwfsix +  paeel +  jclntv +  qjdou 
 var(c( dzwfsix, paeel, jclntv, qjdou ))
zkgdtp <- 90 
woypnze <- 91 
zfyqzi <- 67 
uwxge <- 94 
kigyp <- 65 + hjqsj +  ijtis +  engpslr 
 var(c( hjqsj, ijtis, engpslr ))
pjhpwh <- 35 
vdlbwyo <- 24 
qmyfg <- 114 + weipv +  woypnze 
 var(c( weipv, woypnze ))
heisrl <- 36 
hbptt <- 138 + uezwz +  gbckkh +  gvdlvgg 
 var(c( uezwz, gbckkh, gvdlvgg ))
cneno <- 190 + uwjobru +  jeadxth 
 var(c( uwjobru, jeadxth ))
wkrgvo <- 95 
dcvfkk <- 111 + uahcz +  xsxnizc 
 var(c( uahcz, xsxnizc ))
lfgqgs <- 187 + vbjpz +  hlbotz +  boalcdm 
 var(c( vbjpz, hlbotz, boalcdm ))
xbnmg <- 48 + aesiz +  tqelui 
 var(c( aesiz, tqelui ))
blkewbk <- 64 + jlsryou +  idokb +  avecfoq +  ozyqew 
 var(c( jlsryou, idokb, avecfoq, ozyqew ))
wgsgaj <- 17 
fzwzr <- 12 
wnowypb <- 83 
mifwbmm <- 114 + tfezflv +  qxrspl 
 var(c( tfezflv, qxrspl ))
cbavt <- 161 + qvrvcu +  uhuees +  pggjo +  jajdx 
 var(c( qvrvcu, uhuees, pggjo, jajdx ))
mahlbe <- 6 
xeihyok <- 81 + uunpyb +  wfgzr 
 var(c( uunpyb, wfgzr ))
nnncfl <- 335 + sgytd +  nreuwh 
 var(c( sgytd, nreuwh ))
lkjhmiq <- 25 
teygd <- 170 + llpbev +  nrxyk 
 var(c( llpbev, nrxyk ))
alhghui <- 279 + qbhhtw +  ltytvpb 
 var(c( qbhhtw, ltytvpb ))
khyca <- 84 
atrzkqf <- 95 
olvztw <- 21 
iqjqlfq <- 59 + gmzcjo +  rrzyzq 
 var(c( gmzcjo, rrzyzq ))
oxoul <- 12 + fkzxb +  chcjswk 
 var(c( fkzxb, chcjswk ))
buznuv <- 26 
pbzvjg <- 77 
agztuz <- 53 
zfqmic <- 67 + lkuwf +  pkipy 
 var(c( lkuwf, pkipy ))
talptwp <- 77 
tvxvtce <- 52 + ujgeml +  jpmrs +  lpsaav 
 var(c( ujgeml, jpmrs, lpsaav ))
lfliwg <- 65 
pqrpkmd <- 34 
maftysp <- 53 
oqeam <- 128 + qjzss +  gwuixj 
 var(c( qjzss, gwuixj ))
tlpfhx <- 147 + heisrl +  hskhxx 
 var(c( heisrl, hskhxx ))
tqtut <- 74 + ykitsi +  wxvjyn 
 var(c( ykitsi, wxvjyn ))
epcjds <- 26 
ankrgg <- 55 
thftfut <- 61 
wkkmvcn <- 7 
iytvm <- 15 + juoaena +  qebqgxd +  uwxge 
 var(c( juoaena, qebqgxd, uwxge ))
ruedk <- 146 + kgwftu +  xlacmu 
 var(c( kgwftu, xlacmu ))
ydpemw <- 3192 + ymkgx +  nhtfal 
 var(c( ymkgx, nhtfal ))
steez <- 38 
tkjofj <- 60 + mjqkaqs +  lnfjj +  nakli 
 var(c( mjqkaqs, lnfjj, nakli ))
nasbs <- 70 
valby <- 106 + vgfhpwb +  xtnvc 
 var(c( vgfhpwb, xtnvc ))
admjq <- 40 + qvaayfl +  uqynzy 
 var(c( qvaayfl, uqynzy ))
zqpxn <- 196 + himyt +  zvtfm +  xpotznx 
 var(c( himyt, zvtfm, xpotznx ))
kkmax <- 37 
dfmjy <- 399 + isxysg +  dxlpr 
 var(c( isxysg, dxlpr ))
xcfxvd <- 92 
ejekn <- 98 + kemkz +  souqntk 
 var(c( kemkz, souqntk ))
hsvmept <- 30 
xowkw <- 41 
reomle <- 191 + jacmj +  amcixd 
 var(c( jacmj, amcixd ))
gzocw <- 110 + qkjdjpe +  yxxdal 
 var(c( qkjdjpe, yxxdal ))
ewvvb <- 83 
nlfdps <- 87 
kvvkbhi <- 55 
ncuzpin <- 39 
sybvdj <- 83 
pvyrkg <- 51 + yytmc +  clqfbvi +  tmkli 
 var(c( yytmc, clqfbvi, tmkli ))
qwotm <- 169 + pqrpkmd +  whrinsh 
 var(c( pqrpkmd, whrinsh ))
jfhzgp <- 67 
wyryqvy <- 90 
sgovqnf <- 69 + kzsoasg +  dnbejx 
 var(c( kzsoasg, dnbejx ))
ydffzgu <- 380 + olvztw +  zfipnh 
 var(c( olvztw, zfipnh ))
kerjk <- 84 
grxdbx <- 34 
eovobtl <- 149 + hzuyh +  iftcnc 
 var(c( hzuyh, iftcnc ))
dyomhm <- 40 + npxexv +  xbccghy +  jncjxue +  nadbegy 
 var(c( npxexv, xbccghy, jncjxue, nadbegy ))
temju <- 53 + qpzszp +  cqlvkj 
 var(c( qpzszp, cqlvkj ))
eitdb <- 197 + trvtkqj +  kwnqw 
 var(c( trvtkqj, kwnqw ))
kzjtb <- 78 + gacseeh +  lycfbo +  aifpxqv 
 var(c( gacseeh, lycfbo, aifpxqv ))
ewgadin <- 39 
yznmd <- 62 + uvkwg +  mncivb 
 var(c( uvkwg, mncivb ))
mieqogh <- 19 
qbaosu <- 314 + hshafy +  ukgbdva 
 var(c( hshafy, ukgbdva ))
izxpf <- 39 + ewobo +  ntezb +  sbmqxu +  tlpfggk 
 var(c( ewobo, ntezb, sbmqxu, tlpfggk ))
tmzxl <- 23 
lcmlbj <- 66 
vtccvv <- 69 
nachvlp <- 20 
iwwqvoa <- 24 
sfwzqk <- 54 + rkuzg +  ewvvb 
 var(c( rkuzg, ewvvb ))
devsz <- 38 
vmmtsr <- 33 
ebwubc <- 70 
dmvyuf <- 35 
vtdbov <- 50 
qiksfr <- 44 
wowxv <- 92 
xquhgig <- 190 + rqhfxty +  maftysp 
 var(c( rqhfxty, maftysp ))
igdkn <- 43 + dgtxjuj +  jxyakll +  kptdem 
 var(c( dgtxjuj, jxyakll, kptdem ))
wfsaglg <- 54 + nlfdps +  tdkdiqe 
 var(c( nlfdps, tdkdiqe ))
lqbmor <- 230 + rkjfx +  vmmtsr 
 var(c( rkjfx, vmmtsr ))
qvxanzu <- 262 + nxhfgx +  vvzmdr +  inddon +  ybosi 
 var(c( nxhfgx, vvzmdr, inddon, ybosi ))
inlft <- 36 
xjcxk <- 24 
kcfgbv <- 73 + fdeowqq +  pbzvjg +  tzrqwt 
 var(c( fdeowqq, pbzvjg, tzrqwt ))
ytrsmfx <- 66 
mieynp <- 21 
ihvefep <- 110 + ffzzzc +  ckzaj +  zfyqzi 
 var(c( ffzzzc, ckzaj, zfyqzi ))
dmkfnm <- 30 
ntxcpwl <- 36 + kerjk +  nggubzz +  utapoe 
 var(c( kerjk, nggubzz, utapoe ))
ijkpfv <- 218 + rxcgwd +  steez 
 var(c( rxcgwd, steez ))
hrdggo <- 152 + agztuz +  mehuxr 
 var(c( agztuz, mehuxr ))
keurs <- 18 
mbwaxn <- 240 + nachvlp +  prcxksh 
 var(c( nachvlp, prcxksh ))
wqojj <- 22 
xiiqh <- 98 
zrnlkl <- 63 + dyfzj +  xviqup +  atrzkqf +  ulbute 
 var(c( dyfzj, xviqup, atrzkqf, ulbute ))
irokors <- 39 + xiiqh +  paoxff 
 var(c( xiiqh, paoxff ))
uwjkolt <- 52 + tvxvtce +  igdkn +  rteyxil +  qvxanzu +  qllime 
 var(c( tvxvtce, igdkn, rteyxil, qvxanzu, qllime ))
rvqopbu <- 67 
owukmji <- 195 
ccxff <- 53 + cgveph +  tzbmy 
 var(c( cgveph, tzbmy ))
jfqddfd <- 64 + ewgadin +  fyxttac +  gvtcy 
 var(c( ewgadin, fyxttac, gvtcy ))
noibcw <- 95 
pgaiut <- 90 + vtccvv +  mgcnohi 
 var(c( vtccvv, mgcnohi ))
rpjfkh <- 55 
zesfmu <- 37 
xvhgvw <- 36 
axzszf <- 167 + iiuqtqw +  vktjns 
 var(c( iiuqtqw, vktjns ))
oonpy <- 15 + wyryqvy +  zkgdtp 
 var(c( wyryqvy, zkgdtp ))
syogw <- 36 + mevujmj +  wpphwjv 
 var(c( mevujmj, wpphwjv ))
unkkoph <- 119 + myhnuu +  lcmlbj 
 var(c( myhnuu, lcmlbj ))
syvza <- 99 
icxzod <- 15 
dhisacj <- 73 + rndvcf +  oqmxlu +  ncuzpin 
 var(c( rndvcf, oqmxlu, ncuzpin ))
vjobpzv <- 24 
cokvree <- 96 
claolm <- 312 + ndmkbul +  rpjfkh 
 var(c( ndmkbul, rpjfkh ))
pyhof <- 21 
mgqil <- 66 + noibcw +  nptqxtu 
 var(c( noibcw, nptqxtu ))
yzcez <- 62 + rlxto +  vkaxa 
 var(c( rlxto, vkaxa ))
byedxex <- 180 + gsaqlad +  ebwubc 
 var(c( gsaqlad, ebwubc ))
dsjcqpn <- 133 + slufzy +  gbbpqj +  jxawk 
 var(c( slufzy, gbbpqj, jxawk ))
mxwgu <- 310 + yeyxf +  qdnghwn 
 var(c( yeyxf, qdnghwn ))
ybixf <- 299 + xvhgvw +  ooczbeh +  ksbwlir +  pkzwosb 
 var(c( xvhgvw, ooczbeh, ksbwlir, pkzwosb ))
pralyx <- 191 + dmkfnm +  xcmzg 
 var(c( dmkfnm, xcmzg ))
zpxgfq <- 118 + tqplqn +  fysxwmd 
 var(c( tqplqn, fysxwmd ))
hgkmod <- 75 
yvlajs <- 35 
hkltivy <- 26 
vptdap <- 23 
lsbvzaf <- 75 + ibqrz +  cokvree +  volompq +  zgxbjj 
 var(c( ibqrz, cokvree, volompq, zgxbjj ))
pbdafpz <- 236 + wjdeth +  zacwh +  mxpgmte +  wjgjhc 
 var(c( wjdeth, zacwh, mxpgmte, wjgjhc ))
ipnvjje <- 80 
oyyctv <- 34 + ipnvjje +  tvbgr 
 var(c( ipnvjje, tvbgr ))
fewgct <- 90 
qejgaf <- 25 
wotus <- 119 + dndrnc +  tebkmhi +  qejgaf 
 var(c( dndrnc, tebkmhi, qejgaf ))
hawhqex <- 71 
diqdy <- 183 + gackyrq +  vcjfe +  lcnsf 
 var(c( gackyrq, vcjfe, lcnsf ))
benqzp <- 92 + wvmtyez +  gmveehb 
 var(c( wvmtyez, gmveehb ))
mrpji <- 61 
kzrgfb <- 92 
jtlqxy <- 66 
mpvgy <- 35 
jfsgqq <- 100 + crslvc +  esized 
 var(c( crslvc, esized ))
mcucysr <- 49 
tcbgf <- 28 
jjsyexg <- 38 
kksizqf <- 44 
qakiz <- 37 
xoigv <- 10 
iuehdsr <- 21 
jzgvab <- 9 
odxpup <- 89 
veabcj <- 67 
obmcr <- 89 
ictsk <- 122 + devsz +  jjsyexg 
 var(c( devsz, jjsyexg ))
gixezcy <- 37 
oxzhz <- 86 + bcbftl +  hjzgzoe 
 var(c( bcbftl, hjzgzoe ))
ufrhi <- 68 
ldqyt <- 50 
btwmd <- 54 
aftnyad <- 188 + rfqddq +  recxxj 
 var(c( rfqddq, recxxj ))
nszqngg <- 44 
pnqycz <- 43 + gkkit +  khyca +  yudzxoa 
 var(c( gkkit, khyca, yudzxoa ))
hvwyae <- 23 
jrhzmbj <- 45 
pjndde <- 422 + yseezd +  cseckdt 
 var(c( yseezd, cseckdt ))
iqdiez <- 134 + wkkmvcn +  nkzgse 
 var(c( wkkmvcn, nkzgse ))
khzqqwi <- 12 
iohigm <- 248 + icxzod +  mclqrmw +  pjkyydd 
 var(c( icxzod, mclqrmw, pjkyydd ))
hzehjo <- 43 
wgdnyy <- 75 
wuuvnlz <- 206 + pznkdcr +  ucfipvi +  jfhzgp 
 var(c( pznkdcr, ucfipvi, jfhzgp ))
nnyws <- 190 
bbbhx <- 5 
wayzzn <- 50 
fiynyre <- 156 + mieqogh +  zpaivc 
 var(c( mieqogh, zpaivc ))
vqtalpm <- 26 
dporud <- 50 
qcorw <- 83 
htitjjt <- 203 + nswcez +  msqxo 
 var(c( nswcez, msqxo ))
xsygbj <- 130 + vdlbwyo +  ciquv +  msdcz +  kplvk 
 var(c( vdlbwyo, ciquv, msdcz, kplvk ))
cdfmatu <- 80 
ssnpxj <- 56 
mjqzvuu <- 195 + hvwyae +  tmzxl 
 var(c( hvwyae, tmzxl ))
wzczb <- 61 
rsfxf <- 123 + yjnzgl +  glrmew 
 var(c( yjnzgl, glrmew ))
vhzfut <- 96 + obzrtar +  qtzelhq 
 var(c( obzrtar, qtzelhq ))
rloxobx <- 421 + jiwfia +  nxuer +  xfcktep 
 var(c( jiwfia, nxuer, xfcktep ))
fioegdo <- 140 + opzxu +  blyms 
 var(c( opzxu, blyms ))
sbmwazv <- 39 + arrvxtj +  hawhqex 
 var(c( arrvxtj, hawhqex ))
zmdkea <- 154 + bgnfpcr +  hgkmod 
 var(c( bgnfpcr, hgkmod ))
qjkfys <- 353 + bjhxuqc +  cvxqnh 
 var(c( bjhxuqc, cvxqnh ))
kejph <- 98 + alneot +  jjiikvt 
 var(c( alneot, jjiikvt ))
acvzsbe <- 37 
gqvmeg <- 52 
rdorrx <- 56 
qqooh <- 15 
nzawh <- 22 
ywwxi <- 49 
phnuqn <- 44 
pjoqde <- 76 
fgcajp <- 50 
tefdmi <- 44 
cwxspl <- 12 + nnciip +  hbbpp 
 var(c( nnciip, hbbpp ))
lxexhp <- 35 
scnhwcs <- 74 
qyfsc <- 47 
uxjaa <- 131 + coapfqs +  kxrbd +  rkctaq +  xfyqx 
 var(c( coapfqs, kxrbd, rkctaq, xfyqx ))
eoxwp <- 133 + zxygmaq +  xadjes +  nulax +  rpwdig 
 var(c( zxygmaq, xadjes, nulax, rpwdig ))
gjjfx <- 131 + btwmd +  xznhz +  dsezlbe 
 var(c( btwmd, xznhz, dsezlbe ))
itxgz <- 370 + epqowm +  jrhzmbj 
 var(c( epqowm, jrhzmbj ))
cujuai <- 157 + ldqyt +  fgcajp 
 var(c( ldqyt, fgcajp ))
kgphv <- 246 + wvvivob +  sjoov 
 var(c( wvvivob, sjoov ))
lcrbt <- 29 
xadcuw <- 1626 + chgfb +  stofi 
 var(c( chgfb, stofi ))
qhlhfk <- 387 + iwwqvoa +  rmshk +  ftezz 
 var(c( iwwqvoa, rmshk, ftezz ))
jpscxfh <- 46 + qsakd +  exuyuk +  dqdmie +  kjwfeoj 
 var(c( qsakd, exuyuk, dqdmie, kjwfeoj ))
rusndcc <- 39 + xcfxvd +  kvylxr +  vwkegak +  oiuztx 
 var(c( xcfxvd, kvylxr, vwkegak, oiuztx ))
tkmfbk <- 93 + ffhsg +  kynezb 
 var(c( ffhsg, kynezb ))
vlwvot <- 84 
habni <- 27 + xwknab +  xukfcd +  scvhl +  tcbgf 
 var(c( xwknab, xukfcd, scvhl, tcbgf ))
mgzocpe <- 51 + zkwajbd +  mrpji 
 var(c( zkwajbd, mrpji ))
nvrjzg <- 214 + nzfko +  qqmchoq 
 var(c( nzfko, qqmchoq ))
xvwok <- 34 
xvzoy <- 35 
anamgt <- 346 + ywwxi +  rnbvid 
 var(c( ywwxi, rnbvid ))
rywtmt <- 158 + nevrtrs +  sybvdj +  qcorw 
 var(c( nevrtrs, sybvdj, qcorw ))
pwhtvdh <- 40 
tirifqs <- 6489 + ofbqk +  iqdiez +  kspsz +  kejph 
 var(c( ofbqk, iqdiez, kspsz, kejph ))
bjbndbg <- 169 + oqsdzpe +  sbwoye +  yhukqf 
 var(c( oqsdzpe, sbwoye, yhukqf ))
kgbzu <- 180 + wzczb +  bzqjb 
 var(c( wzczb, bzqjb ))
ifxzdfy <- 96 + ssnpxj +  ejhojmh 
 var(c( ssnpxj, ejhojmh ))
onzind <- 245 + wayzzn +  ystgyng 
 var(c( wayzzn, ystgyng ))
jppbvbo <- 929 + ynvesz +  tvwmzal +  mbwaxn 
 var(c( ynvesz, tvwmzal, mbwaxn ))
mvgmf <- 105 + dicalia +  qhkhuow +  phnuqn 
 var(c( dicalia, qhkhuow, phnuqn ))
krnbkr <- 82 + qladcl +  tefdmi +  ocdpr 
 var(c( qladcl, tefdmi, ocdpr ))
dsmjo <- 32 
gpcyq <- 55 
nzcqte <- 140 + wkkdc +  pwhtvdh +  gugaaiq 
 var(c( wkkdc, pwhtvdh, gugaaiq ))
bywbrk <- 80 + rnkbli +  vlwvot 
 var(c( rnkbli, vlwvot ))
empixc <- 142 + hlwesdc +  xjcxk +  fmhfnhv 
 var(c( hlwesdc, xjcxk, fmhfnhv ))
ypqkl <- 64 + qjeuqe +  wowxv 
 var(c( qjeuqe, wowxv ))
rkgoovn <- 104 + qyfsc +  lytrpo 
 var(c( qyfsc, lytrpo ))
praagwp <- 30 
ogowst <- 176 + pdzooh +  xlccnkv 
 var(c( pdzooh, xlccnkv ))
jblwph <- 81 + lxexhp +  vbqpp 
 var(c( lxexhp, vbqpp ))
rxjly <- 133 + rftuf +  hzehjo 
 var(c( rftuf, hzehjo ))
euwab <- 49 + cdfmatu +  wawtrk 
 var(c( cdfmatu, wawtrk ))
nmazby <- 120 + yijwaf +  idmqvse +  dsmjo +  wqilo 
 var(c( yijwaf, idmqvse, dsmjo, wqilo ))
thhakxn <- 103 + pplis +  tykcvi +  mklqpn +  jzgvab 
 var(c( pplis, tykcvi, mklqpn, jzgvab ))
zvbqni <- 105 + qkzpcyy +  qsrxouy 
 var(c( qkzpcyy, qsrxouy ))
lnndami <- 68 
ycuuwxa <- 161 + vehziv +  rvqopbu 
 var(c( vehziv, rvqopbu ))
mgbgb <- 334 + kksizqf +  dofou 
 var(c( kksizqf, dofou ))
gwcdf <- 32 
hfnhd <- 152 + vdcfh +  ufrhi 
 var(c( vdcfh, ufrhi ))
kjjee <- 174 + lltgjo +  gwcdf 
 var(c( lltgjo, gwcdf ))
bbtnj <- 170 + xvwok +  rqyxzn 
 var(c( xvwok, rqyxzn ))
nrgzcge <- 132 + bywbrk +  vazoq +  eaufvn 
 var(c( bywbrk, vazoq, eaufvn ))
wjxgg <- 177 + symdxo +  dsptf +  yaxeloo +  yxrhis 
 var(c( symdxo, dsptf, yaxeloo, yxrhis ))
eavas <- 45 + ytsmc +  mrxdp +  prjpkm +  nzawh 
 var(c( ytsmc, mrxdp, prjpkm, nzawh ))
hdvcz <- 156 + zisdcf +  keurs +  bjops 
 var(c( zisdcf, keurs, bjops ))
wabzx <- 209 + sfazwz +  lnndami 
 var(c( sfazwz, lnndami ))
wnjaw <- 369 + brvxyyg +  hsvmept +  praagwp 
 var(c( brvxyyg, hsvmept, praagwp ))
rnbgz <- 131 + iuehdsr +  xgzluhk 
 var(c( iuehdsr, xgzluhk ))
pyiexlj <- 92 + oozbxqb +  kzrgfb +  klege +  csfjt 
 var(c( oozbxqb, kzrgfb, klege, csfjt ))
vylzag <- 33 
greav <- 20 
ylbftp <- 82 + epcjds +  gsyeaes 
 var(c( epcjds, gsyeaes ))
ndikjqv <- 22 
ckfefl <- 115 + khzqqwi +  fzwzr 
 var(c( khzqqwi, fzwzr ))
oypvhy <- 140 + yvlajs +  pjhpwh 
 var(c( yvlajs, pjhpwh ))
mhnnmp <- 8 + qbaosu +  pjndde +  anamgt +  bcesxba +  blkewbk 
 var(c( qbaosu, pjndde, anamgt, bcesxba, blkewbk ))
qhbrmur <- 1450 + kcfgbv +  dyomhm +  nofpmh 
 var(c( kcfgbv, dyomhm, nofpmh ))
cywvxf <- 279 + greav +  kybtla 
 var(c( greav, kybtla ))
igytdcp <- 160 + vtdbov +  dporud 
 var(c( vtdbov, dporud ))
gumncu <- 24 
zlhpt <- 74 + zvbqni +  iytvm +  htitjjt +  aqxaecv 
 var(c( zvbqni, iytvm, htitjjt, aqxaecv ))
auvuxv <- 242 + wqojj +  ndikjqv 
 var(c( wqojj, ndikjqv ))
exsmez <- 1357 + pralyx +  tkmfbk +  unkkoph 
 var(c( pralyx, tkmfbk, unkkoph ))
tzimies <- 65 
ddgnyfo <- 79 
ccgjf <- 99 
olawkuc <- 79 + earmfdp +  wnowypb 
 var(c( earmfdp, wnowypb ))
oybbzr <- 1007 + qstlv +  nayvvrh +  izxpf +  habni +  ckfefl +  fjdlcbh +  thhakxn 
 var(c( qstlv, nayvvrh, izxpf, habni, ckfefl, fjdlcbh, thhakxn ))
lyzaead <- 939 + jfqddfd +  ucpkpfb +  sbmwazv 
 var(c( jfqddfd, ucpkpfb, sbmwazv ))
cdjxwb <- 99 + buznuv +  hkltivy 
 var(c( buznuv, hkltivy ))
ucilvib <- 40 + ddgnyfo +  koxzeku 
 var(c( ddgnyfo, koxzeku ))
gtbgp <- 234 + ilsgqdw +  qakiz +  zesfmu 
 var(c( ilsgqdw, qakiz, zesfmu ))
lzojty <- 214 
fnnykbh <- 91 + nxqwag +  tpcbt +  elysbb +  xlaqpg 
 var(c( nxqwag, tpcbt, elysbb, xlaqpg ))
lxwkpqh <- 325 + xoigv +  fqsjaz 
 var(c( xoigv, fqsjaz ))
xohxq <- 1457 + czbydf +  wgdnyy 
 var(c( czbydf, wgdnyy ))
gkwyfz <- 49 
rxjnad <- 80 
qftwuv <- 36 
csrpy <- 62 + zmoxc +  ccgjf +  umrqldb +  gdamb 
 var(c( zmoxc, ccgjf, umrqldb, gdamb ))
oglcdgs <- 140 + dmvyuf +  xvzoy 
 var(c( dmvyuf, xvzoy ))
uxolsws <- 25 
gzwrv <- 212 + efqjc +  gasejp +  qqooh 
 var(c( efqjc, gasejp, qqooh ))
vrfme <- 20 
ozdnanh <- 34 
ezlfnhs <- 62 + qftwuv +  ynlzggk 
 var(c( qftwuv, ynlzggk ))
pygrhr <- 144 + rdorrx +  pjecv 
 var(c( rdorrx, pjecv ))
icush <- 877 + rsfxf +  zfqmic +  axzszf +  gzwrv +  cujuai 
 var(c( rsfxf, zfqmic, axzszf, gzwrv, cujuai ))
jagajrh <- 93 
kviiza <- 70 
anggtqq <- 86 
ntlrkw <- 36 + gkwyfz +  acjedny +  mcucysr 
 var(c( gkwyfz, acjedny, mcucysr ))
emkgc <- 142 + lcrbt +  pcvfgti 
 var(c( lcrbt, pcvfgti ))
fsdoohr <- 64 
egkiqcp <- 37 + rxjnad +  psetts 
 var(c( rxjnad, psetts ))
vtcxw <- 860 + ezlfnhs +  aizdt +  ylbftp 
 var(c( ezlfnhs, aizdt, ylbftp ))
hgwbg <- 89 
zicvok <- 81 
ejdxula <- 92 + kviiza +  uqmaluk +  nasbs 
 var(c( kviiza, uqmaluk, nasbs ))
jykseco <- 166 + xawzbz +  fsdoohr 
 var(c( xawzbz, fsdoohr ))
fqvbdq <- 96 + tzimies +  jduij 
 var(c( tzimies, jduij ))
gugxitg <- 553 + trlcj +  olawkuc +  xeihyok 
 var(c( trlcj, olawkuc, xeihyok ))
agjpn <- 8265 + ocgbpez +  ejdxula +  kgbzu 
 var(c( ocgbpez, ejdxula, kgbzu ))
angpy <- 62 + fitha +  itxgz +  pyiexlj +  wqshh +  rloxobx 
 var(c( fitha, itxgz, pyiexlj, wqshh, rloxobx ))
mkgafux <- 688 + admjq +  tqtut +  emkgc 
 var(c( admjq, tqtut, emkgc ))
vxslvq <- 37 + jmvcgo +  ioglbe 
 var(c( jmvcgo, ioglbe ))
dydcy <- 76 + drnki +  uxolsws +  lrcstqr 
 var(c( drnki, uxolsws, lrcstqr ))
uwxjkq <- 905 + jblwph +  cdjxwb +  dydcy 
 var(c( jblwph, cdjxwb, dydcy ))
xaden <- 108 + jagajrh +  fucxhz 
 var(c( jagajrh, fucxhz ))
chyun <- 94 
hycgb <- 237 + zmciyk +  ysbhsn +  vxslvq 
 var(c( zmciyk, ysbhsn, vxslvq ))
vukpdsr <- 162 + gumncu +  mlzmimg +  vjobpzv 
 var(c( gumncu, mlzmimg, vjobpzv ))
kzjozhi <- 90 
ixncawy <- 28 
dzfon <- 61 + wksjlix +  zicvok +  utiwem 
 var(c( wksjlix, zicvok, utiwem ))
jzoxv <- 52 
kgxfc <- 30 + kzjozhi +  dtqgdc 
 var(c( kzjozhi, dtqgdc ))
nxbec <- 96 
ercxmph <- 318 + ixncawy +  zhrdezs 
 var(c( ixncawy, zhrdezs ))
bmukspb <- 106 + jzoxv +  gqvmeg 
 var(c( jzoxv, gqvmeg ))
cqjsqh <- 360 + bzmqg +  jpscxfh +  ercxmph 
 var(c( bzmqg, jpscxfh, ercxmph ))
htaszsq <- 87 
hppxlu <- 174 + lkjhmiq +  kntwy +  auhnyn 
 var(c( lkjhmiq, kntwy, auhnyn ))
nhvyso <- 77 
mogwbda <- 60 
yswonja <- 195 + gachtig +  bjadb +  gtiopzn 
 var(c( gachtig, bjadb, gtiopzn ))
piizso <- 59 
gyirkcj <- 138 + inlft +  vkihn 
 var(c( inlft, vkihn ))
tduxvx <- 326 + iktget +  vylzag +  picnycy +  lfdery 
 var(c( iktget, vylzag, picnycy, lfdery ))
hiuviy <- 59 
zaargo <- 30 + ytrsmfx +  jtlqxy 
 var(c( ytrsmfx, jtlqxy ))
nfywcyp <- 62 
npttye <- 42 + wkrgvo +  jhydv +  enmjipp +  hnyjyga 
 var(c( wkrgvo, jhydv, enmjipp, hnyjyga ))
jtxjj <- 120 + hiuviy +  sadyx 
 var(c( hiuviy, sadyx ))
aceyee <- 219 + gixezcy +  zqwqcnq 
 var(c( gixezcy, zqwqcnq ))
bwgimi <- 242 + eohjrft +  ankrgg +  kvvkbhi 
 var(c( eohjrft, ankrgg, kvvkbhi ))
bzjxil <- 23 
tlqjsb <- 994 + odxpup +  znagag +  obmcr 
 var(c( odxpup, znagag, obmcr ))
pxxabj <- 33 + wgupfl +  lpdhdvy +  lfliwg +  fklfgd 
 var(c( wgupfl, lpdhdvy, lfliwg, fklfgd ))
ygqxl <- 16 + nhvyso +  talptwp +  qiznqk +  ugzmeb 
 var(c( nhvyso, talptwp, qiznqk, ugzmeb ))
hocazx <- 818 + bmukspb +  mgbwxw +  gyirkcj +  kgxfc 
 var(c( bmukspb, mgbwxw, gyirkcj, kgxfc ))
pmolg <- 166 + uuunqk +  uolqt 
 var(c( uuunqk, uolqt ))
nzuswm <- 80 + ojzogs +  qfopwo 
 var(c( ojzogs, qfopwo ))
ubvttg <- 796 + wuuvnlz +  donzb +  bwgimi +  rywtmt +  rusndcc +  ceagctk 
 var(c( wuuvnlz, donzb, bwgimi, rywtmt, rusndcc, ceagctk ))
ohcoit <- 6 
dfofyir <- 236 + ddswb +  grxdbx 
 var(c( ddswb, grxdbx ))
gkfdg <- 226 + ohcoit +  jkoir 
 var(c( ohcoit, jkoir ))
byahs <- 65 
iffxdk <- 87 
wzbewii <- 157 + myetb +  dhwakp +  vqtalpm 
 var(c( myetb, dhwakp, vqtalpm ))
nqjydpi <- 44 + rpgzp +  nszqngg +  jthpj +  qiksfr 
 var(c( rpgzp, nszqngg, jthpj, qiksfr ))
qtxmjwm <- 7 
ckudisa <- 78 
opskc <- 35 
qhkniqm <- 96 
qqmvd <- 14 
swgjx <- 6 
tzngdx <- 486 + pvyrkg +  hppxlu +  eovobtl +  diqdy 
 var(c( pvyrkg, hppxlu, eovobtl, diqdy ))
ktjdpt <- 23 + tdmirp +  qhpvuzi 
 var(c( tdmirp, qhpvuzi ))
invra <- 180 + sxdngcn +  mahlbe +  swgjx 
 var(c( sxdngcn, mahlbe, swgjx ))
nkwwia <- 1405 + dsjcqpn +  irokors +  wzbewii 
 var(c( dsjcqpn, irokors, wzbewii ))
oausk <- 31 
pjujovx <- 96 + ayuma +  bmmzqxb 
 var(c( ayuma, bmmzqxb ))
gqqsfsr <- 78 + ckudisa +  kwhgtnk 
 var(c( ckudisa, kwhgtnk ))
ibqmynm <- 252 + oglcdgs +  tkjofj +  upurae +  oypvhy +  bzfkt +  hdvcz +  cfwxyl 
 var(c( oglcdgs, tkjofj, upurae, oypvhy, bzfkt, hdvcz, cfwxyl ))
rkerea <- 87 
mxemqqb <- 267 + iuuouds +  qqmvd 
 var(c( iuuouds, qqmvd ))
ewqcktv <- 1364 + hfnhd +  amqvr +  ntxcpwl 
 var(c( hfnhd, amqvr, ntxcpwl ))
mzutqi <- 571 + qdfvw +  mxwgu +  yswonja 
 var(c( qdfvw, mxwgu, yswonja ))
bggovyb <- 144 + supjxi +  mieynp +  wltax +  pyhof 
 var(c( supjxi, mieynp, wltax, pyhof ))
veuzidj <- 49 + kmhug +  yrolq +  csfzl 
 var(c( kmhug, yrolq, csfzl ))
xbxxe <- 173 + bzjxil +  amklj 
 var(c( bzjxil, amklj ))
sflonq <- 190 + gppwsbp +  mpvgy 
 var(c( gppwsbp, mpvgy ))
aemkbp <- 35 + dwrtbo +  rkerea +  iffxdk 
 var(c( dwrtbo, rkerea, iffxdk ))
dpngy <- 17 
hxswr <- 74 + bixxpzd +  htaszsq 
 var(c( bixxpzd, htaszsq ))
rziopi <- 57 + flgpg +  nqjydpi +  xbnmg +  xxdcgas +  vhzfut +  sfwzqk +  kirnt 
 var(c( flgpg, nqjydpi, xbnmg, xxdcgas, vhzfut, sfwzqk, kirnt ))
ejoxsy <- 38 + lxwkpqh +  zdelqb +  xyang +  gtbgp +  onzind +  wabzx +  kzjtb 
 var(c( lxwkpqh, zdelqb, xyang, gtbgp, onzind, wabzx, kzjtb ))
agthzo <- 15 
xyvnk <- 32 
crchwpb <- 36 
xnxudsh <- 78 + pjlxc +  nfywcyp 
 var(c( pjlxc, nfywcyp ))
vahmn <- 49 + twyhx +  rztur +  pchfpqp +  anggtqq 
 var(c( twyhx, rztur, pchfpqp, anggtqq ))
girxc <- 147 + zylotoh +  qtxmjwm 
 var(c( zylotoh, qtxmjwm ))
fkttd <- 141 + hbhfdlh +  hgwbg 
 var(c( hbhfdlh, hgwbg ))
oqjqu <- 67 + xbxxe +  yjwvnhl +  vmsoc +  tlpfhx +  rxjly +  dyazidi 
 var(c( xbxxe, yjwvnhl, vmsoc, tlpfhx, rxjly, dyazidi ))
gcswiam <- 1183 + vahmn +  cwfydbh +  cjjytkr 
 var(c( vahmn, cwfydbh, cjjytkr ))
ofkvpq <- 500 + auvuxv +  stqmwk +  dptekr 
 var(c( auvuxv, stqmwk, dptekr ))
nlgymop <- 139 + tymkixg +  crchwpb 
 var(c( tymkixg, crchwpb ))
fcgkmq <- 68 + thftfut +  svtqdbd 
 var(c( thftfut, svtqdbd ))
uhrdeol <- 171 + qinne +  opskc +  zxmwy +  dgqfrv 
 var(c( qinne, opskc, zxmwy, dgqfrv ))
rarkt <- 1387 + mjqzvuu +  nzhbfc +  veuzidj 
 var(c( mjqzvuu, nzhbfc, veuzidj ))
pqiuhyp <- 52 + xbtjl +  jykseco +  aftnyad +  xaden +  urdgg +  ijkpfv +  ejekn 
 var(c( xbtjl, jykseco, aftnyad, xaden, urdgg, ijkpfv, ejekn ))
cqvdpy <- 62 + yzcez +  nvrjzg +  sebvkw +  pgaiut +  bggovyb +  wfsaglg +  teygd 
 var(c( yzcez, nvrjzg, sebvkw, pgaiut, bggovyb, wfsaglg, teygd ))
uzljl <- 710 + pmolg +  benqzp +  khfzis 
 var(c( pmolg, benqzp, khfzis ))
bobvssz <- 25 + srvqt +  alhghui +  wnjaw +  qhlhfk +  lsbvzaf +  qjkfys +  nnncfl 
 var(c( srvqt, alhghui, wnjaw, qhlhfk, lsbvzaf, qjkfys, nnncfl ))
zqbpucs <- 568 + dfmjy +  fnnykbh +  gqeiy +  fifams 
 var(c( dfmjy, fnnykbh, gqeiy, fifams ))
domway <- 2464 + kgphv +  pjujovx +  hrdggo 
 var(c( kgphv, pjujovx, hrdggo ))
rtwlo <- 78 + wgbvle +  npttye +  claolm +  ydffzgu +  mgbgb +  jwjuth +  mulvwn 
 var(c( wgbvle, npttye, claolm, ydffzgu, mgbgb, jwjuth, mulvwn ))
tymmo <- 22 + uxjaa +  nlgymop +  dcvfkk +  yzjhmqq +  iqjqlfq +  faopk 
 var(c( uxjaa, nlgymop, dcvfkk, yzjhmqq, iqjqlfq, faopk ))
rreidgp <- 168 + vfupn +  xyvnk 
 var(c( vfupn, xyvnk ))
hpziqqg <- 637 + ypqkl +  nmazby +  hxswr 
 var(c( ypqkl, nmazby, hxswr ))
mjrktxe <- 50 + eoxwp +  jgsmitm +  isrgei +  rnbgz +  kymbvpp +  ktjdpt +  mgzocpe 
 var(c( eoxwp, jgsmitm, isrgei, rnbgz, kymbvpp, ktjdpt, mgzocpe ))
jazqb <- 7378 + xadcuw +  hocazx +  cqvdpy 
 var(c( xadcuw, hocazx, cqvdpy ))
slzpby <- 136 + nxxlbnt +  ozdnanh +  zppujp 
 var(c( nxxlbnt, ozdnanh, zppujp ))
onjqcrh <- 67 + lkdyq +  tptqtpq +  dmidur +  klkwsa 
 var(c( lkdyq, tptqtpq, dmidur, klkwsa ))
xrtzs <- 96 + ictsk +  invra +  rkgoovn +  gzocw +  hbptt +  ucilvib +  cnffplo 
 var(c( ictsk, invra, rkgoovn, gzocw, hbptt, ucilvib, cnffplo ))
bujufb <- 148 + tduxvx +  lxzvkpk +  jllqymn +  csrpy 
 var(c( tduxvx, lxzvkpk, jllqymn, csrpy ))
mcrhiy <- 6559 + mkgafux +  gugxitg +  tymmo 
 var(c( mkgafux, gugxitg, tymmo ))
hsdym <- 340 + onjqcrh +  eitdb +  ycuuwxa +  mxemqqb +  pnqycz +  wexncjl 
 var(c( onjqcrh, eitdb, ycuuwxa, mxemqqb, pnqycz, wexncjl ))
hrrld <- 1802 + pqiuhyp +  nkwwia +  exsmez +  rarkt +  hsdym 
 var(c( pqiuhyp, nkwwia, exsmez, rarkt, hsdym ))
vyinb <- 154 + fzyim +  vrfme +  hjbfr 
 var(c( fzyim, vrfme, hjbfr ))
wscqe <- 14 + ezzpr +  chyun 
 var(c( ezzpr, chyun ))
njsjq <- 273 + vptdap +  eteyvz 
 var(c( vptdap, eteyvz ))
vwnvlh <- 1534 + krnbkr +  egizyey +  empixc +  lzojty +  vyinb +  zeolg +  qxnvoh 
 var(c( krnbkr, egizyey, empixc, lzojty, vyinb, zeolg, qxnvoh ))
udgssh <- 5 
xahvbm <- 222 + udgssh +  bbbhx 
 var(c( udgssh, bbbhx ))
foznw <- 146 + dpngy +  wgsgaj +  xboqv 
 var(c( dpngy, wgsgaj, xboqv ))
zyqssjb <- 13 + olebdv +  byahs +  xajzwdh 
 var(c( olebdv, byahs, xajzwdh ))
ptatzl <- 314 + sfdclf +  ibbxb +  slzpby +  kjjee +  bbtnj +  gkfdg +  nzuswm 
 var(c( sfdclf, ibbxb, slzpby, kjjee, bbtnj, gkfdg, nzuswm ))
cmsqe <- 12 + njsjq +  zqpxn +  pulql +  fkttd +  wozakri 
 var(c( njsjq, zqpxn, pulql, fkttd, wozakri ))
bvaoy <- 114 + mogwbda +  rodntl 
 var(c( mogwbda, rodntl ))
rswnjx <- 57 
vfrylx <- 10 + qhkniqm +  bmcwbsi 
 var(c( qhkniqm, bmcwbsi ))
bbbvmxo <- 21 + rceksnv +  uhnlse +  rswnjx +  ogcmm 
 var(c( rceksnv, uhnlse, rswnjx, ogcmm ))
codkti <- 71 + tqbtvqc +  oausk 
 var(c( tqbtvqc, oausk ))
bcadp <- 91 
dzuqljn <- 7 + spnllx +  fpeiss +  hjzkjuw +  pjoqde 
 var(c( spnllx, fpeiss, hjzkjuw, pjoqde ))
igcmpn <- 55 + bcadp +  sxjlr 
 var(c( bcadp, sxjlr ))
wugia <- 2689 + temju +  xkymbld +  ntlrkw 
 var(c( temju, xkymbld, ntlrkw ))
qsdyww <- 55 
puvmdie <- 41 
umjcr <- 59 
iffluz <- 515 + bbbvmxo +  reomle +  vspqi 
 var(c( bbbvmxo, reomle, vspqi ))
dcero <- 34 + ssvfbfs +  nxbec 
 var(c( ssvfbfs, nxbec ))
vutdji <- 31 + bsvgagj +  puvmdie +  liujmc +  xowkw 
 var(c( bsvgagj, puvmdie, liujmc, xowkw ))
pmopji <- 128 + gpcyq +  qsdyww 
 var(c( gpcyq, qsdyww ))
gozhrsf <- 762 + ifxzdfy +  uuxpvxf +  zyqssjb 
 var(c( ifxzdfy, uuxpvxf, zyqssjb ))
wjrhp <- 992 + nyouzhg +  ccxff +  qmifv +  oonpy +  vutdji +  owukmji 
 var(c( nyouzhg, ccxff, qmifv, oonpy, vutdji, owukmji ))
yuiyjpc <- 862 + eavas +  lxttas +  codkti 
 var(c( eavas, lxttas, codkti ))
krgtrdn <- 53 
yyecrv <- 59 
aicjff <- 67 
ayjse <- 75 + veabcj +  aicjff 
 var(c( veabcj, aicjff ))
nrgxdmx <- 1352 + mlalrpt +  nqfqgj +  cuscn +  ogowst +  vfrylx 
 var(c( mlalrpt, nqfqgj, cuscn, ogowst, vfrylx ))
fjticim <- 35 
fnslg <- 52 
tokyi <- 79 + yapfoe +  yyecrv 
 var(c( yapfoe, yyecrv ))
rcluk <- 156 + fjticim +  mbbuqoi 
 var(c( fjticim, mbbuqoi ))
thkjz <- 437 + insleik +  ayjse +  sgovqnf +  wjxgg +  euwab 
 var(c( insleik, ayjse, sgovqnf, wjxgg, euwab ))
hwgwzv <- 57 + qqqbf +  fnslg 
 var(c( qqqbf, fnslg ))
ibeis <- 54 
mohey <- 86 + tbmyue +  ibeis 
 var(c( tbmyue, ibeis ))
xnnbqju <- 90 + xzsmw +  petvm +  xyxidwf 
 var(c( xzsmw, petvm, xyxidwf ))
vxiwms <- 99 
ropdc <- 226 + usevjww +  agthzo 
 var(c( usevjww, agthzo ))
jefsr <- 477 + fqvbdq +  xsygbj +  rcluk +  dcero +  lfgqgs 
 var(c( fqvbdq, xsygbj, rcluk, dcero, lfgqgs ))
mlvhqpe <- 91 + qsstg +  tnrrui +  utsnke +  gxtfu 
 var(c( qsstg, tnrrui, utsnke, gxtfu ))
bdlnohy <- 584 + dodrbx +  qwotm +  cbavt +  mvgmf +  igcmpn 
 var(c( dodrbx, qwotm, cbavt, mvgmf, igcmpn ))
yxbaldn <- 122 + vxiwms +  syvza 
 var(c( vxiwms, syvza ))
zwopz <- 60 + byedxex +  yxbaldn +  hnkvw +  lvkeqm +  kigyp +  ruedk 
 var(c( byedxex, yxbaldn, hnkvw, lvkeqm, kigyp, ruedk ))
bxqqvue <- 5668 + zqbpucs +  mhnnmp +  ewqcktv 
 var(c( zqbpucs, mhnnmp, ewqcktv ))
qxugs <- 31 + iqmpc +  lrtzi +  krgtrdn 
 var(c( iqmpc, lrtzi, krgtrdn ))
pyxdkm <- 872 + xnnbqju +  oxzhz +  zaargo 
 var(c( xnnbqju, oxzhz, zaargo ))
vtjtyp <- 152 + dhisacj +  zpxgfq +  oqeam +  fcgkmq +  nnyws +  qxugs +  oxoul 
 var(c( dhisacj, zpxgfq, oqeam, fcgkmq, nnyws, qxugs, oxoul ))
lcitl <- 1114 + nlwakg +  hwgwzv +  girxc 
 var(c( nlwakg, hwgwzv, girxc ))
vqinar <- 2904 + qhbrmur +  nrgxdmx +  angpy +  gcswiam 
 var(c( qhbrmur, nrgxdmx, angpy, gcswiam ))
rbtco <- 1110 + dfofyir +  qnhbmzp +  dzfon +  zmdkea +  valby +  mabcqcb +  mifwbmm 
 var(c( dfofyir, qnhbmzp, dzfon, zmdkea, valby, mabcqcb, mifwbmm ))
ylnobx <- 5204 + rziopi +  mzutqi +  lcitl 
 var(c( rziopi, mzutqi, lcitl ))
fgsmzi <- 525 + dzuqljn +  ihvefep +  uhrdeol +  gdvuw 
 var(c( dzuqljn, ihvefep, uhrdeol, gdvuw ))
fmpvv <- 74 
sjbfmfa <- 80 + fewgct +  jupzgjm 
 var(c( fewgct, jupzgjm ))
fivkqxx <- 121 + acvzsbe +  kkmax +  qcmwsvm 
 var(c( acvzsbe, kkmax, qcmwsvm ))
umtxk <- 2942 + jyazs +  scnhwcs +  fmpvv +  aahqvqy 
 var(c( jyazs, scnhwcs, fmpvv, aahqvqy ))
ndticx <- 54 + rbtco +  bobvssz +  umtxk +  domway +  wugia +  ubvttg +  ydpemw 
 var(c( rbtco, bobvssz, umtxk, domway, wugia, ubvttg, ydpemw ))
jynmg <- 1140 + wotus +  mohey +  fiynyre 
 var(c( wotus, mohey, fiynyre ))
wzimze <- 4005 + uwjkolt +  jynmg +  ibqmynm 
 var(c( uwjkolt, jynmg, ibqmynm ))
vhkodl <- 221 + otlvq +  fioegdo +  rreidgp +  fivkqxx +  xahvbm 
 var(c( otlvq, fioegdo, rreidgp, fivkqxx, xahvbm ))
uqals <- 78 + ropdc +  pbdafpz +  mgqil +  pygrhr +  eourjv 
 var(c( ropdc, pbdafpz, mgqil, pygrhr, eourjv ))
abxglwt <- 191 + pmopji +  jtxjj +  asevk +  yznmd +  jfsgqq 
 var(c( pmopji, jtxjj, asevk, yznmd, jfsgqq ))
zildb <- 116 + umjcr +  pinqp +  piizso 
 var(c( umjcr, pinqp, piizso ))
sjmhuxn <- 11 + aceyee +  iohigm +  zildb +  vrmqym +  pxxabj +  gjjfx 
 var(c( aceyee, iohigm, zildb, vrmqym, pxxabj, gjjfx ))
fonrb <- 7531 + xohxq +  cmsqe +  jefsr 
 var(c( xohxq, cmsqe, jefsr ))
phqpwkw <- 1192 + egkiqcp +  ecjxs +  foznw +  tokyi 
 var(c( egkiqcp, ecjxs, foznw, tokyi ))
tespoy <- 5 + bdlnohy +  jppbvbo +  fgsmzi +  sjmhuxn 
 var(c( bdlnohy, jppbvbo, fgsmzi, sjmhuxn ))
xtnmps <- 95 + ptatzl +  bujufb +  oybbzr +  zwopz +  phqpwkw 
 var(c( ptatzl, bujufb, oybbzr, zwopz, phqpwkw ))
oxypy <- 1992 + sflonq +  igytdcp +  nzcqte +  sjbfmfa 
 var(c( sflonq, igytdcp, nzcqte, sjbfmfa ))
ikzap <- 5388 + mjrktxe +  tlqjsb +  yuiyjpc 
 var(c( mjrktxe, tlqjsb, yuiyjpc ))
exrud <- 176 + hpziqqg +  abxglwt +  gozhrsf +  oqjqu +  vhkodl 
 var(c( hpziqqg, abxglwt, gozhrsf, oqjqu, vhkodl ))
hrjlzxr <- 1124 + ybixf +  qmshniw +  zrnlkl 
 var(c( ybixf, qmshniw, zrnlkl ))
bwlit <- 20 
hnjszii <- 284 + zkdwz +  bwlit 
 var(c( zkdwz, bwlit ))
rhqfb <- 179 + okurp +  gpwbcg +  invqsb 
 var(c( okurp, gpwbcg, invqsb ))
pekohn <- 65 
nxmhned <- 166 + pekohn +  xnszvs 
 var(c( pekohn, xnszvs ))
glcsinj <- 90 + aemkbp +  yxxgd +  fhqhu +  xquhgig +  nxmhned +  lqbmor +  qmyfg 
 var(c( aemkbp, yxxgd, fhqhu, xquhgig, nxmhned, lqbmor, qmyfg ))
mflijpx <- 75 + oxypy +  rtwlo +  vwnvlh 
 var(c( oxypy, rtwlo, vwnvlh ))
mxnaq <- 3937 + glcsinj +  icush +  wjrhp 
 var(c( glcsinj, icush, wjrhp ))
rbzmniw <- 52959 + tespoy +  tirifqs +  exrud 
 var(c( tespoy, tirifqs, exrud ))
mlmhe <- 4563 + uwxjkq +  uqals +  ofkvpq +  pyxdkm 
 var(c( uwxjkq, uqals, ofkvpq, pyxdkm ))
zqamk <- 37518 + mflijpx +  ikzap +  wzimze +  agjpn 
 var(c( mflijpx, ikzap, wzimze, agjpn ))
tigvdj <- 44217 + xtnmps +  ylnobx +  mlmhe 
 var(c( xtnmps, ylnobx, mlmhe ))
xqwfgy <- 1496 + cywvxf +  bjbndbg +  mlvhqpe 
 var(c( cywvxf, bjbndbg, mlvhqpe ))
cvqbwem <- 92 + wjceo +  gqqsfsr +  vukpdsr +  cneno +  bvaoy 
 var(c( wjceo, gqqsfsr, vukpdsr, cneno, bvaoy ))
gewpohj <- 16410 + cvqbwem +  uzljl +  zlhpt +  vtcxw +  iffluz 
 var(c( cvqbwem, uzljl, zlhpt, vtcxw, iffluz ))
cjgtra <- 900 + nfkfex +  rhqfb +  oyyctv 
 var(c( nfkfex, rhqfb, oyyctv ))
ktasnia <- 49 + thkjz +  cjgtra +  tzngdx +  cqjsqh +  vtjtyp +  lyzaead +  xrtzs 
 var(c( thkjz, cjgtra, tzngdx, cqjsqh, vtjtyp, lyzaead, xrtzs ))
wdbmakv <- 42933 + ktasnia +  mxnaq +  mcrhiy 
 var(c( ktasnia, mxnaq, mcrhiy ))
wxgoew <- 1481 + hnjszii +  ygqxl +  xhkwrs 
 var(c( hnjszii, ygqxl, xhkwrs ))
dcpzwz <- 87 + wxgoew +  xqwfgy +  hrjlzxr +  wapknj +  ejoxsy 
 var(c( wxgoew, xqwfgy, hrjlzxr, wapknj, ejoxsy ))
hvecp <- 90 + hrrld +  jazqb +  fonrb +  dcpzwz +  vqinar +  bxqqvue 
 var(c( hrrld, jazqb, fonrb, dcpzwz, vqinar, bxqqvue ))
 
 # missed above, so added
bqyqwn <- 68 + wscqe + cwxspl + syogw + xnxudsh
 
xflhoc <- 20092 + bqyqwn +  hycgb +  nrgzcge 
 var(c( bqyqwn, hycgb, nrgzcge ))
jfofam <- 6042 + gewpohj +  ndticx +  xflhoc 
 var(c( gewpohj, ndticx, xflhoc ))
eqgvf <- 9 + wdbmakv +  zqamk +  hvecp +  rbzmniw +  jfofam +  tigvdj 
 var(c( wdbmakv, zqamk, hvecp, rbzmniw, jfofam, tigvdj ))

```
```{r}
c( wdbmakv, zqamk, hvecp, rbzmniw, jfofam, tigvdj )
# rbzmniw depends on exrud
c( tespoy, tirifqs, exrud )
# exrud depends on gozhrsf
c( hpziqqg, abxglwt, gozhrsf, oqjqu, vhkodl )
c( ifxzdfy, uuxpvxf, zyqssjb ) #is balanced
# gozhrsf should be 762 - 5
```

