---
title: "Day 9"
author: "Derek Sollberger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
```

--- Day 9: Stream Processing ---

A large stream blocks your path. According to the locals, it's not safe to cross the stream at the moment because it's full of garbage. You look down at the stream; rather than water, you discover that it's a stream of characters.

You sit for a while and record part of the stream (your puzzle input). The characters represent groups - sequences that begin with { and end with }. Within a group, there are zero or more other things, separated by commas: either another group or garbage. Since groups can contain other groups, a } only closes the most-recently-opened unclosed group - that is, they are nestable. Your puzzle input represents a single, large group which itself contains many smaller ones.

Sometimes, instead of a group, you will find garbage. Garbage begins with < and ends with >. Between those angle brackets, almost any character can appear, including { and }. Within garbage, < has no special meaning.

In a futile attempt to clean up the garbage, some program has canceled some of the characters within it using !: inside garbage, any character that comes after ! should be ignored, including <, >, and even another !.

You don't see any characters that deviate from these rules. Outside garbage, you only find well-formed groups, and garbage always terminates according to the rules above.

Here are some self-contained pieces of garbage:

    <>, empty garbage.
    <random characters>, garbage containing random characters.
    <<<<>, because the extra < are ignored.
    <{!>}>, because the first > is canceled.
    <!!>, because the second ! is canceled, allowing the > to terminate the garbage.
    <!!!>>, because the second ! and the first > are canceled.
    <{o"i!a,<{i<a>, which ends at the first >.

Here are some examples of whole streams and the number of groups they contain:

    {}, 1 group.
    {{{}}}, 3 groups.
    {{},{}}, also 3 groups.
    {{{},{},{{}}}}, 6 groups.
    {<{},{},{{}}>}, 1 group (which itself contains garbage).
    {<a>,<a>,<a>,<a>}, 1 group.
    {{<a>},{<a>},{<a>},{<a>}}, 5 groups.
    {{<!>},{<!>},{<!>},{<a>}}, 2 groups (since all but the last > are canceled).

Your goal is to find the total score for all groups in your input. Each group is assigned a score which is one more than the score of the group that immediately contains it. (The outermost group gets a score of 1.)

    {}, score of 1.
    {{{}}}, score of 1 + 2 + 3 = 6.
    {{},{}}, score of 1 + 2 + 2 = 5.
    {{{},{},{{}}}}, score of 1 + 2 + 3 + 3 + 3 + 4 = 16.
    {<a>,<a>,<a>,<a>}, score of 1.
    {{<ab>},{<ab>},{<ab>},{<ab>}}, score of 1 + 2 + 2 + 2 + 2 = 9.
    {{<!!>},{<!!>},{<!!>},{<!!>}}, score of 1 + 2 + 2 + 2 + 2 = 9.
    {{<a!>},{<a!>},{<a!>},{<ab>}}, score of 1 + 2 = 3.

What is the total score for all groups in your input?

```{r}
stream_score <- function(stream){
  depth <- 0
  garbage_mode <- FALSE
  skip_next <- FALSE
  N <- str_length(stream)
  our_score <- 0
  for(i in 1:N){
    this_character <- str_sub(stream, i, i)
    if(skip_next){
      skip_next <- FALSE
    }else{
      if(this_character == "!"){ #escaping
        skip_next <- TRUE
      }else{
        if(this_character == "<"){
          garbage_mode <- TRUE
        }else if(this_character == ">"){
          garbage_mode <- FALSE
        }else{
          
          if(garbage_mode){
            # do nothing
          }else{
            if(this_character == "{"){
              depth <- depth + 1
            }else if(this_character == "}"){
              our_score <- our_score + depth
              depth <- depth - 1
            }
          }
        }
      }
    }
  }
  our_score #return
}

# test cases
stream_score("{}")
stream_score("{{{}}}")
stream_score("{{},{}}")
stream_score("{{{},{},{{}}}}")
stream_score("{<a>,<a>,<a>,<a>}")
stream_score("{{<ab>},{<ab>},{<ab>},{<ab>}}")
stream_score("{{<!!>},{<!!>},{<!!>},{<!!>}}")
stream_score("{{<a!>},{<a!>},{<a!>},{<ab>}}")
```

```{r}
puzzle_input <- readLines("20171209input.txt")
print("The total score for all groups in our input")
print(paste("is", stream_score(puzzle_input)))
```

--- Part Two ---

Now, you're ready to remove the garbage.

To prove you've removed it, you need to count all of the characters within the garbage. The leading and trailing < and > don't count, nor do any canceled characters or the ! doing the canceling.

    <>, 0 characters.
    <random characters>, 17 characters.
    <<<<>, 3 characters.
    <{!>}>, 2 characters.
    <!!>, 0 characters.
    <!!!>>, 0 characters.
    <{o"i!a,<{i<a>, 10 characters.

How many non-canceled characters are within the garbage in your puzzle input?

```{r}
garbage_count <- function(stream){
  depth <- 0
  garbage_mode <- FALSE
  skip_next <- FALSE
  N <- str_length(stream)
  our_score <- 0
  for(i in 1:N){
    this_character <- str_sub(stream, i, i)
    if(skip_next){
      skip_next <- FALSE
    }else{
      if(this_character == "!"){ #escaping
        skip_next <- TRUE
      }else{
        if(this_character == "<" & !garbage_mode){
          garbage_mode <- TRUE
        }else if(this_character == "<" & garbage_mode){
          our_score <- our_score + 1
        }else if(this_character == ">"){
          garbage_mode <- FALSE
        }else{
          
          if(garbage_mode){
            # CHANGE WAS HERE
            our_score <- our_score + 1
          }else{
            if(this_character == "{"){
              depth <- depth + 1
            }else if(this_character == "}"){
              # our_score <- our_score + depth
              depth <- depth - 1
            }
          }
        }
      }
    }
  }
  our_score #return
}

# test cases
garbage_count("<>")
garbage_count("<random characters>")
garbage_count("<<<<>")
garbage_count("<{!>}>")
garbage_count("<!!>")
garbage_count("<!!!>>")
# garbage_count("<{o'i!a,<{i<a>") #typo in prompt??
```

```{r}
print("The total garbage count")
print(paste("was",
            garbage_count(puzzle_input),
            "characters."))
```

