---
title: "Day 13"
author: "Derek Sollberger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

--- Day 13: A Maze of Twisty Little Cubicles ---

You arrive at the first floor of this new building to discover a much less welcoming environment than the shiny atrium of the last one. Instead, you are in a maze of twisty little cubicles, all alike.

Every location in this area is addressed by a pair of non-negative integers (x,y). Each such coordinate is either a wall or an open space. You can't move diagonally. The cube maze starts at 0,0 and seems to extend infinitely toward positive x and y; negative values are invalid, as they represent a location outside the building. You are in a small waiting area at 1,1.

While it seems chaotic, a nearby morale-boosting poster explains, the layout is actually quite logical. You can determine whether a given x,y coordinate will be a wall or an open space using a simple system:

    Find x*x + 3*x + 2*x*y + y + y*y.
    Add the office designer's favorite number (your puzzle input).
    Find the binary representation of that sum; count the number of bits that are 1.
        If the number of bits that are 1 is even, it's an open space.
        If the number of bits that are 1 is odd, it's a wall.

For example, if the office designer's favorite number were 10, drawing walls as # and open spaces as ., the corner of the building containing 0,0 would look like this:

  0123456789
0 .#.####.##
1 ..#..#...#
2 #....##...
3 ###.#.###.
4 .##..#..#.
5 ..##....#.
6 #...##.###

Now, suppose you wanted to reach 7,4. The shortest route you could take is marked as O:

  0123456789
0 .#.####.##
1 .O#..#...#
2 #OOO.##...
3 ###O#.###.
4 .##OO#OO#.
5 ..##OOO.#.
6 #...##.###

Thus, reaching 7,4 would take a minimum of 11 steps (starting from your current location, 1,1).

What is the fewest number of steps required for you to reach 31,39?

```{r}
wall_maker <- function(x, y, puzzle_input = 10){
  z <- x*x + 3*x + 2*x*y + y + y*y
  z <- z + puzzle_input
  z_bits <- as.integer(intToBits(z))
  if((sum(z_bits == 1) %% 2) == 0){
    this_space <- "."
  }else{
    this_space <- "#"
  }
  this_space # return
}

# test cases
wall_maker(0,1)
wall_maker(1,0)
wall_maker(9,5)
wall_maker(9,6)
```

```{r}
maze_maker <- function(num_rows, num_cols, puzzle_input = 10){
  # intentionally making a walls on top and left sides of maze
  M <- matrix("#", nrow = num_rows, ncol = num_cols)
  for(i in 0:(num_rows - 2)){
    for(j in 0:(num_cols - 2)){
      M[i+2,j+2] <- wall_maker(j,i,puzzle_input)
    }
  }
  M[3,3] <- 0 #starting location
  M # return
}

# test case
maze_maker(8, 11, 10)
```

```{r}
path_finder <- function(maze){
  num_rows <- nrow(maze)
  num_cols <- ncol(maze)
  for(k in 1:max(num_cols, num_rows)){
    for(i in 2:(num_rows - 1)){
      for(j in 2:(num_cols - 1)){
        this_entry <- maze[i,j]
        if(this_entry != "#" & this_entry != "."){
          this_entry <- as.integer(this_entry)
          
          #check up, left, right, and down
          if(maze[i-1,j] == "."){
            maze[i-1,j] <- this_entry + 1
          }
          if(maze[i,j-1] == "."){
            maze[i,j-1] <- this_entry + 1
          }
          if(maze[i,j+1] == "."){
            maze[i,j+1] <- this_entry + 1
          }
          if(maze[i+1,j] == "."){
            maze[i+1,j] <- this_entry + 1
          }
        }
      }
    }
  }
  maze # return
}

# test it out
path_finder(maze_maker(8, 11, 10))
```

```{r}
# example
# M <- path_finder(maze_maker(8, 11, 10))
# print(paste("It would take",
#             M[4+2, 7+2],
#             "steps to reach the destination."))
# real puzzle
M <- path_finder(maze_maker(50, 50, 1364))
print(paste("It would take",
            M[39+2, 31+2],
            "steps to reach the destination."))
```
--- Part Two ---

How many locations (distinct x,y coordinates, including your starting location) can you reach in at most 50 steps?

```{r}
M <- path_finder(maze_maker(100, 100, 1364))
M[M == "#"] <- 99
M[M == "."] <- 99
M <- as.integer(M)
print(paste("Our range includes",
            sum(M <= 50),
            "locations."))
```

