---
title: "Day 12"
author: "Derek Sollberger"
date: "11/5/2021"
output: html_document
---

```{r}
library("rjson")
library("tidyverse")
```


--- Day 12: JSAbacusFramework.io ---

Santa's Accounting-Elves need help balancing the books after a recent order. Unfortunately, their accounting software uses a peculiar storage format. That's where you come in.

They have a JSON document which contains a variety of things: arrays ([1,2,3]), objects ({"a":1, "b":2}), numbers, and strings. Your first job is to simply find all of the numbers throughout the document and add them together.

For example:

    [1,2,3] and {"a":2,"b":4} both have a sum of 6.
    [[[3]]] and {"a":{"b":4},"c":-1} both have a sum of 3.
    {"a":[-1,1]} and [-1,{"a":1}] both have a sum of 0.
    [] and {} both have a sum of 0.

You will not encounter any strings containing numbers.

What is the sum of all numbers in the document?

```{r}
# does str_extract() work here?
# https://stackoverflow.com/questions/53818952/extracting-digits-after-only-with-negative-number
sum(as.numeric(str_extract_all("[1,2,3]", "-?[0-9]+")[[1]]))
sum(as.numeric(str_extract_all("{\"a\":2,\"b\":4}", "-?[0-9]+")[[1]]))
sum(as.numeric(str_extract_all("{\"a\":{\"b\":4},\"c\":-1}", "-?[0-9]+")[[1]]))
sum(as.numeric(str_extract_all("{\"a\":[-1,1]}", "-?[0-9]+")[[1]]))
sum(as.numeric(str_extract_all("[-1,{\"a\":1}]", "-?[0-9]+")[[1]]))
sum(as.numeric(str_extract_all("[]", "-?[0-9]+")[[1]]))
sum(as.numeric(str_extract_all("{}", "-?[0-9]+")[[1]]))
```
```{r}
# helper function
sum_all_numbers <- function(this_string){
  sum(as.numeric(str_extract_all(this_string, "-?[0-9]+")[[1]]))
}
```

```{r}
# load data
todays_data <- readLines("20151212input.txt")
```
```{r}
sum_all_numbers(todays_data)
```
--- Part Two ---

Uh oh - the Accounting-Elves have realized that they double-counted everything red.

Ignore any object (and all of its children) which has any property with the value "red". Do this only for objects ({...}), not arrays ([...]).

    [1,2,3] still has a sum of 6.
    [1,{"c":"red","b":2},3] now has a sum of 4, because the middle object is ignored.
    {"d":"red","e":[1,2,3,4],"f":5} now has a sum of 0, because the entire structure is ignored.
    [1,"red",5] has a sum of 6, because "red" in an array has no effect.

```{r}
todays_new_data <- todays_data
object_list <- str_extract_all(todays_data, "(?<=\\{).+?(?=\\})")[[1]]
```

```{r}
# exploring
test_string <- "[1,{\"c\":\"red\",\"b\":2},3]"
test_object <- str_extract_all(test_string, "(?<=\\{).+?(?=\\})")[[1]]
str_detect(test_object, "red")
# str_extract(test_object, "\\{(.*?)\\}")
# str_replace(test_object, "\"c\":\"red\",\"b\":2", "")
str_replace(test_string, test_object, "")
sum_all_numbers("[1,{},3]")

bracket_locations <- str_locate_all(test_string, "(?<=\\{).+?(?=\\})")[[1]]
str_sub(test_string, bracket_locations[1, "start"], 
        bracket_locations[1, "end"]) <- ""
```





Ran into NESTED objects, so I need to be clever

```{r}
# get string location for braces
curly_braces_locations <- 
  data.frame(str_locate_all(todays_data, "(?<=\\{).+?(?=\\})")[[1]])
curly_braces_locations <- curly_braces_locations %>%
  mutate(object_length = end - start) %>%
  arrange(object_length)
```

```{r, eval = FALSE}
# will need string locations
object_df <- data.frame(object_list)
colnames(object_df) <- c("object")
object_df <- object_df %>%
  mutate(object_length = str_length(object)) %>%
  arrange(object_length)
```


```{r, eval = FALSE}
# early draft
n <- length(object_list)
for(i in 1:n){
  if(str_detect(object_list[i], "red")){
    todays_new_data <- str_replace(todays_new_data, object_list[i], "")
  }
}
```

```{r}
# todays_new_data <- test_object
todays_new_data <- todays_data
n <- length(object_list)
for(i in 1:n){
  curly_braces_locations <- 
    data.frame(str_locate_all(todays_new_data, "(?<=\\{).+?(?=\\})")[[1]])
  curly_braces_locations <- curly_braces_locations %>%
    mutate(object_length = end - start) %>%
    arrange(desc(object_length))
    # arrange(object_length)
  
  # print(nrow(curly_braces_locations))
  # print(paste("red attributes:",
              # str_count(todays_new_data, "red")))
  
  for(j in 1:nrow(curly_braces_locations)){
    this_object <- str_sub(todays_new_data, 
                          curly_braces_locations[j, "start"], 
                          curly_braces_locations[j, "end"])
    if(str_detect(this_object, "red")){
      # todays_new_data <- str_replace(todays_new_data,
      #                                fixed(this_object),
      #                                "")
      str_sub(todays_new_data,
              curly_braces_locations[j, "start"] - 1,
              curly_braces_locations[j, "end"] + 1) <- ""
      break
    }
  }
  # print(str_length(todays_new_data))
}  
```


```{r}
write_lines(todays_new_data, "20151212output.txt")
```


```{r}
str_length(todays_data)
str_length(todays_new_data)
```
```{r}
sum_all_numbers(todays_new_data)
#90360 was too low
#92350 was too low
```

---

```{r}
todays_data_json <- fromJSON(file = "20151212input.json")
```

```{r}
# print(todays_data)
# json_df <- data.frame(todays_data_json)
```

